---
title: "1.2 Prilaminary DD with cyprus"
author: "TG"
date: "2023-01-09"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
in this script I used beal on 0/1 data

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(rfishbase)
library(Hmisc)
library(vegan)
library(mobr)
library(viridis)
library(plotrix)

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\Initial DD plots with cyprus"
```


```{r}

medata_raw <- readRDS("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/cyprus fix/medata_from_git_after_cyp.Rds") # the original data (not really the row-row one)


medata<-medata_raw # the verstion ill play with


zero_obs<- medata %>% filter(sp.n == 0)
```

kornati217015286494381202NA

kornati217615472824388942NA

kornati216815286494381202NA

kornati216915286494381202NA

gdor164534866923241672NA

brijuni210313817044485674NA



convert to wide


```{r}


medata<-medata %>% select(1:20)

medata<-medata %>% group_by(across(1:19)) %>% summarise(amount = sum(sp.n))

medata_wide<-spread(medata,species,amount,fill = 0)

not_unique<-medata_wide %>% group_by(unique_trans_id) %>% summarise(n = n()) %>% filter(n>1)


```


convert to presence/absence

```{r}

medata_wide[,19:153]<-decostand(medata_wide[,19:153],method = "pa") 

```


crate different sp matrix that contains the beal index value for each species

Beals smoothing replaces each entry in the community data with a probability of a target species occurring in that particular site, based on the joint occurrences of the target species with the species that actually occur in the site. 

* should i calculate beals index sepertly for each contry?

```{r}
beals <- medata_wide 

beals[,19:153] <- beals(beals[,19:153],type = 0,include = FALSE) 


```

```{r}
boxplot(unlist(beals[,19:153]) ~ unlist(medata_wide[,19:153]), xlab="Presence", ylab="Beals")


```



see for each species if it was present or absent form a site/year along side with its Beals probability  

```{r}
g_beals <-gather(beals,"species","beal",19:ncol(beals))
g_beals<-g_beals %>% arrange(site,trans)

g_pa<-gather(medata_wide,"species","pa",19:ncol(medata_wide))
g_pa<-g_pa %>% arrange(site,trans)

dark_diversity<-left_join(g_beals,g_pa) 

dark_diversity<-dark_diversity %>%
  mutate(unique_id =paste(site,trans,sep = " "),.after = trans) %>%
  mutate(region = country,.after = country)

```


set a threshold for being included in dark diversity -  the mean beal probability for each species **present** in each country in the data


```{r}

threshold <- dark_diversity %>%
  filter(pa >0) %>% # present species
  group_by(species,country) %>%
  summarise(avg_beal_pr =mean(beal)) # mean beal value for the species when present

dark_diversity<- left_join(dark_diversity,threshold)

dark_diversity$avg_plus_five<-dark_diversity$avg_beal_pr+0.05

dark_diversity$dark_sp<-ifelse(dark_diversity$pa == "0" & dark_diversity$beal > dark_diversity$avg_plus_five,
                               "dark",
                               "normal")

exotic<- medata_raw %>% select(species,exotic) %>% distinct(.keep_all = T)



dark_diversity<-left_join(dark_diversity,exotic,by = "species")

```

lets see what species are in the dd assamlege

```{r}

only_dark <-dark_diversity %>% filter(dark_sp=="dark")


summary(only_dark$protection)
summary(only_dark$enforcement)
summary.factor(only_dark$species)

only_dark

```


```{r}

dark_diversity<-dark_diversity %>% drop_na(avg_beal_pr) # remove rows that 


#rmove species that wasnt observed and we didint expect tham to be there

dark_diversity$remove = ifelse(dark_diversity$pa ==0 & dark_diversity$dark_sp=="normal",
                               "rm","keep")


dark_diversity<- dark_diversity %>% filter(remove =="keep") %>% select(-remove)

dark_diversity$dark_sp[dark_diversity$dark_sp=="normal"]<-"observed"

dark_diversity<-dark_diversity %>% relocate(dark_sp,.after = species)

dark_diversity$protection[dark_diversity$protection=="TRUE"]<-"MPA"
dark_diversity$protection[dark_diversity$protection=="FALSE"]<-"outside MPA"


```

plot the mean richness, DD and Total by protection levels

```{r}

enf<-dark_diversity %>%
  group_by(enforcement,dark_sp,unique_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_id)


enf<-spread(enf,dark_sp,richness,fill = 0)

enf$total<-enf$dark + enf$observed

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_id)

enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


enf %>%  na.omit() %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9))+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean richness per transect")

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))

#ggsave(path = plot_dir, filename = "mean dd and richness by enforcement.png")

enf %>%  na.omit() %>% filter(type == "dark") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()

#ggsave(path = plot_dir, filename = "mean dd by enforcement.png")
```


plot the mean richness, DD and Total by protection levels

```{r}

protection<-dark_diversity %>%
  group_by(protection,dark_sp,unique_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_id)


protection<-spread(protection,dark_sp,richness,fill = 0)

protection$total<-protection$dark + protection$observed

protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_id)

protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9))+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean richness per transect")

protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()


#ggsave(path = plot_dir, filename = "mean dd and richness by mpa.png")

protection %>%  na.omit() %>% filter(type == "dark") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()

#ggsave(path = plot_dir, filename = "mean dd by mpa.png")

protection %>%  na.omit() %>% filter(type == "observed") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill ="#F7D54A")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean richness per transect") +theme_classic()

#ggsave(path = plot_dir, filename = "mean richness by mpa.png")

```