---
title: "2.2 complexity analyses"
author: "TG"
date: "2023-01-03"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(rfishbase)
library(Hmisc)
library(vegan)
library(mobr)
library(viridis)
library(plotrix)
library(mltools)

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\Initial DD plots\\complexity_1"
```


```{r}

comp_data_raw <-read.csv("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/data/Mediterranean_data_COMPLEX.csv")

lab_fish_data_raw <- read.csv("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/data/Med_analyses_data.csv")

comp_data_raw$X<-NULL
lab_fish_data_raw$X<-NULL
```

unite data frames

```{r}

colnames(comp_data_raw)
colnames(lab_fish_data_raw)


all_data<-left_join(lab_fish_data_raw,comp_data_raw)


```

keep one observer only

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(all_data$id)) {
  
  data <- all_data %>% filter(id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)
rm(data,filter_data)
```



```{r}

one_observer <-one_observer %>% select(-right_observer,-left_observer,-ID,-length,-new_a,-new_b, -new_id,-observer_num,-distance,-direction,-campaign) %>% relocate(species,.after = earth_movers) %>%
  relocate(abundance,.after = species)

wide_data<- one_observer %>% group_by(across(1:38)) %>% summarise("total_abundance"= sum(abundance))

wide_data<-spread(wide_data,species,total_abundance,fill = 0)



```

add quantiles to some complexity indeces

```{r}
wide_data$q_vr<-ntile(wide_data$VerRelief,4)
wide_data$q_curve<-ntile(wide_data$Curvature,4)
wide_data$q_freeS<-ntile(wide_data$FreeSpace,4)
wide_data$q_rug<-ntile(wide_data$Rugosity,4)

wide_data<- wide_data %>% relocate(c(77:80),.after = earth_movers)

```


calculate the dark diversity

abundance based beal index:

```{r}
beals <- wide_data 

beals[,42:80] <- beals(beals[,42:80],type = 1,include = FALSE) 
```

see for each species if it was present or absent form a site/year along side with its Beals probability  

```{r}
g_beals <-gather(beals,"species","beal",42:ncol(beals))
g_beals<-g_beals %>% arrange(date,location,id)

g_pa<-gather(wide_data,"species","pa",42:ncol(wide_data))
g_pa<-g_pa %>% arrange(date,location,id)

dark_diversity<-left_join(g_beals,g_pa) 


```

set a threshold for being included in dark diversity -  the mean beal probability for each species **present** in each country in the data


```{r}

threshold <- dark_diversity %>%
  filter(pa > 0) %>% # present species
  group_by(species) %>%
  summarise(avg_beal_pr =mean(beal)) # mean beal value for the species when present

dark_diversity<- left_join(dark_diversity,threshold)

dark_diversity$avg_plus_five<-dark_diversity$avg_beal_pr+0.05

dark_diversity$dark_sp<-ifelse(dark_diversity$pa == "0" & dark_diversity$beal > dark_diversity$avg_plus_five,
                               "dark",
                               "normal")


```


lets see what species are in the dd assamlege

```{r}

only_dark <-dark_diversity %>% filter(dark_sp=="dark")


summary.factor(only_dark$species)



```

```{r}
dark_diversity$remove = ifelse(dark_diversity$pa ==0 & dark_diversity$dark_sp=="normal",
                               "rm","keep")

dark_diversity<- dark_diversity %>% ungroup() %>% filter(remove =="keep") %>% select(-remove,-max_richness)

dark_diversity$dark_sp[dark_diversity$dark_sp=="normal"]<-"observed"

dark_diversity<-dark_diversity %>% relocate(dark_sp,.after = species)


```


```{r}
complexity<-dark_diversity %>%
  group_by(id,reserve,enforcement,VerRelief,Curvature,FreeSpace,Rugosity,dark_sp,
           q_vr,q_curve,q_freeS,q_rug) %>%
  summarise(richness = n_distinct(species)) %>% arrange(id)


complexity<-spread(complexity,dark_sp,richness,fill = 0)

complexity$total<-complexity$dark + complexity$observed



complexity <- gather(complexity,type,richness,12:14)

complexity <- complexity %>% arrange(id) %>% na.omit()

complexity %>% 
ggplot()+aes(x=VerRelief,y=richness,color = type)+
  geom_point()+geom_smooth(method = "lm")+xlab("Vertical relif")

#ggsave(path = plot_dir, filename = "ver_relif.png")

complexity %>% 
ggplot()+aes(x=Curvature,y=richness,color = type)+
  geom_point()+geom_smooth(method = "lm")+xlab("curvature index")

ggsave(path = plot_dir, filename = "curvature.png")


complexity %>% 
ggplot()+aes(x=FreeSpace,y=richness,color = type)+
  geom_point()+geom_smooth(method = "lm")+xlab("Free space")

complexity %>% 
ggplot()+aes(x=Rugosity,y=richness,color = type)+
  geom_point()+geom_smooth(method = "lm")+xlab("rugosity")

```
complexity bins:

vertical relif

```{r}

complex_bins<-dark_diversity %>%
  group_by(id,reserve,enforcement,VerRelief,Curvature,FreeSpace,Rugosity,dark_sp,
           q_vr,q_curve,q_freeS,q_rug) %>%
  summarise(richness = n_distinct(species)) %>% arrange(id)

complex_bins<-spread(complex_bins,dark_sp,richness,fill = 0)

complex_bins$total<-complex_bins$dark + complex_bins$observed

complex_bins <- gather(complex_bins,type,richness,12:14)
complex_bins <- complex_bins %>% arrange(id)

complex_bins_qvr <- complex_bins %>%
  group_by(q_vr,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


complex_bins_qvr %>%   
ggplot()+
       aes(x=q_vr,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9))+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean richness per transect")

complex_bins_qvr %>%  filter(type != "total") %>% 
ggplot()+
       aes(x=q_vr,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  xlab("Complexity (Vertical relif)")


ggsave(path = plot_dir, filename = "complexity_bin_vr.png")
```

Curvature

```{r}
complex_bins_curve <- complex_bins %>%
  group_by(q_curve,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


complex_bins_curve %>%   
ggplot()+
       aes(x=q_curve,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9))+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean richness per transect")

complex_bins_curve %>%  filter(type != "total") %>% 
ggplot()+
       aes(x=q_curve,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  xlab("Complexity (Curvature)")


ggsave(path = plot_dir, filename = "complexity_bin_curveature.png")
```

