---
title: "4.1 DD transect map and stat"
author: "TG"
date: "2023-01-15"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

Get my final data, statistic and map:

```{r message=FALSE, warning=FALSE}

library(lubridate)
library(ggplot2)
library(stringr)
library(rfishbase)
#library(Hmisc)
library(vegan)
library(mobr)
library(viridis)
library(plotrix)
library(sjmisc)
library(sf)
library(ggspatial)
library(rgeos)
library(maptools)
library(dplyr)
library(tidyr)
library(leaflet)

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\4.1 plots"
```


# MEDATA

```{r}

medata_raw <- readRDS("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/cyprus fix/medata_from_git_after_cyp.Rds") # the original data (not really the row-row one)


medata<-medata_raw # the verstion ill play with


```

## Data cleaning

Remove transects that have only 0 and seems like mistakes:

```{r}

zero_obs<- medata %>% filter(sp.n == 0)

transect_to_remove<-c("kornati217015286494381202NA",
                      "kornati216815286494381202NA",
                      "kornati216915286494381202NA",
                      "gdor164534866923241672NA")

medata<-medata %>% filter(!unique_trans_id %in% transect_to_remove)

```

Examine the species list:

```{r}

species<- medata %>% distinct(species)

```

Remove `.spp` species:

```{r}

Species_to_remove <-species$species[which(str_detect(species$species,".spp")==T)]

medata <-medata %>% filter(!species %in% Species_to_remove)

species<- medata %>% distinct(species,family)


```

Remove cryptic species:

Gobiidae, Blenniidae & Tripterygiidae


```{r}

Families_to_remove<- c("Gobiidae","Blenniidae","Tripterygiidae")

medata <-medata %>% filter(!family %in% Families_to_remove)

species<- medata %>% distinct(species,family)

```


## fix cyprus coords

the lat and lon of cyprus

```{r}

medata$lat_n<-ifelse(medata$country=="Cyprus",
                               medata$lon,
                               medata$lat)

medata$lon_n<-ifelse(medata$country=="Cyprus",
                              medata$lat,
                              medata$lon)

medata<-medata %>% dplyr::select(-lat,-lon) %>% 
  rename("lat" = "lat_n","lon" = "lon_n") %>%
  relocate(lat, .after = season) %>% 
  relocate(lon,.before = site)

medata<-medata %>% drop_na(protection)


```

## add banyuls coordinates

```{r}
Banyuls_location <- read_csv("Banyuls location.csv")
Banyuls_location <- Banyuls_location %>%  select(site,Latitude,Longitude)

medata<-left_join(medata,Banyuls_location)

medata$lat<-ifelse(medata$country=="France",medata$Latitude,medata$lat)
medata$lon<-ifelse(medata$country=="France",medata$Longitude,medata$lon)


medata$Latitude<-NULL
medata$Longitude<- NULL


rm(Banyuls_location)
```



## load map

```{r}
med_seas <- st_read("Med_World_Seas_IHO_v3_MarineRegions/Medit_seas_IHO.shp")

```


### which points in which polygon

```{r}
sf_medata<- st_as_sf(medata, coords = c("lon","lat"),crs = "WGS84")

ggplot() +
  geom_sf(data = med_seas) + # plot the polygons first
  geom_sf(data= sf_medata)
```

```{r}
map_n_point <- st_join(sf_medata, med_seas, join = st_within)


ggplot() +
  geom_sf(data = med_seas) + # plot the polygons first
  geom_sf(data= map_n_point,aes(color = NAME))

map_n_point %>% filter(is.na(NAME))

rm(sf_medata)
rm(map_n_point)
```
what happens is that I have many points that fall on land...
I used for loop to fix it: 

 
```{r}
med_seas_sp <- as_Spatial(med_seas)  #create a version of the shp in sp

for (i in 1:nrow(med_seas_sp)) {
  
  region <- med_seas_sp[i,]  #select each region
  region_bf <- gBuffer(region, width = 0.01) #make a buffer around the region
  other_regs <- med_seas_sp[-i,]  #select all other regions
  region_bf_land <- gDifference(region_bf, other_regs) #delete part of the buffer overlapping other regions
  region_bf_land <- spChFIDs(region_bf_land, paste(i))  #indicate ID of each feature
  region_bf_land <-  SpatialPolygonsDataFrame(region_bf_land, region@data) #give back the info stollen by evil functions
  
  if(i == 1){
    med_seas_bf <- region_bf_land #for the first loop
  }else{
    med_seas_bf <- spRbind(med_seas_bf, region_bf_land) #for teh other loopa
  }
  
  print(i)
}

#get the info of see for each point

medata_sp <- as.data.frame(medata) #make a spatialPointsDataFrame object
coordinates(medata_sp) <- ~ lon + lat  #inform lon and lat to transform into spatial object
proj4string(medata_sp) <- proj4string(med_seas_bf) #inform sp points obj coordinate info system
points_regions <- over(medata_sp, med_seas_bf) #extract bassin info for each point
medata_sp$Basin <- points_regions$NAME
medata_sp$colour <- as.integer(as.factor(medata_sp$Basin))
unique(is.na(medata_sp$Basin))

plot(med_seas_bf,col = 'azure', border = NA, bg = 'grey90')
plot(medata_sp, add = T, pch = 19, cex = 0.5, col = medata_sp$colour)
text(med_seas_bf$Longitude, y = med_seas_bf$Latitude,med_seas_bf$NAME) 

```

cool!

now ill add to my non-spatiat medata the basin

```{r}
basin <-medata_sp@data

basin <- basin %>%  dplyr::select(Basin,unique_trans_id) %>% distinct(unique_trans_id,.keep_all = T)

medata<-left_join(medata,basin)

medata<-medata %>% relocate(Basin,.after = country)

```


Create alternative basin division which:

* add the `Ionian Sea` to the `Adriatic Sea`
* add the `Ligurian Sea` to the `Mediterranean Sea - Western Basin`
* separate Israel or Cyprus to the `Levantine Sea`



```{r}
medata<-medata %>% mutate(alt_basin = Basin) %>% relocate(alt_basin,.after = Basin)

medata$alt_basin<-ifelse(medata$Basin == "Ionian Sea",
                         "Adriatic Sea",
                         medata$alt_basin)

medata$alt_basin<-ifelse(medata$Basin == "Ligurian Sea",
                         "Mediterranean Sea - Western Basin",
                         medata$alt_basin)

medata$alt_basin<-ifelse(medata$country %in% c("Israel","Cyprus"),
                         "Levantine Sea",
                         medata$alt_basin)

```

fix some point in turkey that needed to be added to the levent

smaller that lat 29.472579

```{r}

medata$alt_basin<-ifelse(medata$country == "Turkey" & medata$lon > 29.472579,"Levantine Sea",medata$alt_basin)

```


### unite close sites (script 4.2)
```{r}

meta <- medata %>% select(1:20) %>% distinct(unique_trans_id,.keep_all = T)



```

get mean coordiante per site


```{r}
meta_site<- meta %>% group_by(alt_basin,site) %>% summarise(lat = mean(lat),lon = mean(lon))
 

  
```


change all the assecret sites:


```{r}

new_name<-data.frame(site=c("assecret2010193","assecret2010192","assecret2010191","assecret1210192","assecret1210191"),
                     new_site_name = rep("Crete_Monoafitis"))


new_name_2<- data.frame(site=c("assecret1810193","assecret1810192","assecret1810191"),
                     new_site_name = rep("Crete_Paximadia"))

new_name_3<- data.frame(site=c("assecret1410193","assecret1610191","assecret1410192","assecret1410191",
                               "assecret1610192","assecret1610193"),
                     new_site_name = rep("Crete_Plakias"))


new_name_4<- data.frame(site=c("assecret2210191","assecret2210192"),
                     new_site_name = rep("Crete_Krios"))


new_name_5<- data.frame(site=c("assecret1910191","assecret1910193","assecret1910192","assecret1310192",
                               "assecret1310191","assecret1310193"),
                     new_site_name = rep("Crete_Kalydon"))


new_name_6<-data.frame(site = unique(medata$site[medata$country=="France"]),
                       new_site_name = "Banyuls")

new_name_7<-data.frame(site = c("asinara_add","Nisia","montgri"),
                       new_site_name = c("asinara","Kavo Gkreko","medes"))

new_name <- rbind(new_name,new_name_2,new_name_3,new_name_4,new_name_5,new_name_6,new_name_7)



meta_site<-left_join(meta_site,new_name)

meta_site<-meta_site %>% mutate(new_site_name = case_when(is.na(new_site_name) ~ site,
                                   TRUE ~ new_site_name)) %>% 
  
select(site,new_site_name)


medata<-left_join(medata,meta_site)

medata<-medata %>% 
  relocate(new_site_name,.after = site) 

medata$new_site_name<-as.factor(medata$new_site_name)


meta_new <- medata %>% select(1:20) %>% distinct(unique_trans_id,.keep_all = T)


meta_site_new<- meta_new %>% group_by(new_site_name,alt_basin) %>% summarise(lat = mean(lat),lon = mean(lon))
 
```


fix some point in spain that needed to be added to the Balearic

```{r}
medata$alt_basin[medata$new_site_name=="cabrera"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin[medata$new_site_name=="illa_de_l'aire"]<-"Balearic (Iberian Sea)" # one basin for the site

```


change `TRUE` & `FALSE` variables

```{r}

medata$protection<-ifelse(medata$protection==TRUE,"Protected","Not protected")

medata$exotic<-ifelse(medata$exotic==TRUE,"Invasive","Native")

```

# save medata

```{r}

#save(medata,file = "clean_dd_medata.Rdata")

```


## some data statistics

count transects per location

```{r}

transect_per_site<-medata %>%
    group_by(country,new_site_name,protection,enforcement) %>%
  summarise("number_of_trans"= n_distinct(unique_trans_id)) 

```

Get Transect per:

1. country
2. country & protection
3. protecion (Yes/no)
4. Enforcment level
5. basin
6. alternative basin
7. alternative basin + protection

```{r}

#1
transect_per_country<-medata %>% group_by(country) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#2
transect_per_country_pro<-medata %>% group_by(country,protection) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#3

transect_per_protection<-medata %>% group_by(protection) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 
#4

transect_per_enforcment<-medata %>% group_by(enforcement) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#5

transect_per_basin <- medata %>% group_by(Basin) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#6

transect_per_altbasin <- medata %>% group_by(alt_basin) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#7

transect_per_altbasin_protection <- medata %>% group_by(alt_basin,protection) %>% summarise("number_of_trans"= n_distinct(unique_trans_id)) 

#8

transect_per_altbasin_site <- medata %>% group_by(alt_basin) %>% summarise("number_of_trans"= n_distinct(unique_trans_id),"site_number"=n_distinct(new_site_name)) 
```

## Study site map

get only distinct point with different lat and lon:

```{r}
unique_coords<-medata %>% distinct(lat,lon,enforcement,protection,.keep_all = T) %>% dplyr::select(country,site,lat,lon,protection,enforcement)

```

#,position=position_jitter(h=0.12, w=0.12)
```{r}


ggplot(unique_coords) +
  geom_sf(data = med_seas$geometry, mapping = aes(),fill= "azure",color = NA) + ##D4ECFF
  geom_point(data =unique_coords,aes(x = lon, y = lat, color = enforcement), 
             size = 2.5,alpha = 0.9,position=position_jitter(h=0.06, w=0.06)) + 
  xlab("Longitude") + 
  ylab("Latitude") +
   annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"))+
  theme(panel.background = element_rect(fill = "grey90"),##FFFAF7
        text=element_text(size=24, family = "serif"),
        panel.grid.major = element_line(colour = "transparent"))+
    geom_sf_text(data =med_seas, aes(label = med_seas$NAME),
                 family = "serif")+
  scale_color_manual("Protection level",values=c("#C42E1C","#D4B04A","#70A34E","#446441"),
                     labels = c("Not protected","Minimally protected","Partly protection","Fully protected"))
  

ggsave(filename = "med_map_v1.pdf", width=15, height=10)
ggsave(filename = "med_map_v1.png", width=15, height=10)
```


`#C42E1C` - red
`#D4B04A`- yellow
`#70A34E`- light green
`#446441` - dark green

Ill take it to QGIS!

```{r}
color_pal<- colorFactor(palette = c("#C42E1C","#D4B04A","#70A34E","#446441"), domain = unique_coords$enforcement,alpha = 0.8)


leaflet(unique_coords) %>% 
  addTiles() %>% 
  addCircles(lng=~lon,lat=~lat,radius = 50,color = ~color_pal(unique_coords$enforcement),
    stroke = T,label = ~as.character(site))
```



# save medata

```{r}

#save(medata,file = "clean_dd_medata.Rdata")

```

### save data

save the unique coordinates data


```{r}
#save(unique_coords,file = "unique_coords.Rdata")
#write.csv(unique_coords,"unique_coords.csv")

```

# complexity data

```{r}

comp_data_raw <-read.csv("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/data/Mediterranean_data_COMPLEX.csv")

lab_fish_data_raw <- read.csv("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/data/Med_analyses_data.csv")

comp_data_raw$X<-NULL
lab_fish_data_raw$X<-NULL
```

unite data frames

```{r}

colnames(comp_data_raw)
colnames(lab_fish_data_raw)


all_data<-left_join(lab_fish_data_raw,comp_data_raw)

```

keep one observer only

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(all_data$id)) {
  
  data <- all_data %>% filter(id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)
rm(data,filter_data)
```



```{r}

one_observer <-one_observer %>% select(-right_observer,-left_observer,-ID,-new_a,-new_b, -new_id,-observer_num,-distance,-direction,-campaign) %>% relocate(species,.after = earth_movers) %>%
  relocate(abundance,.after = species) %>% 
  relocate(length,.before = abundance)


one_observer<-one_observer %>%  select(-visibility,-time,-max_richness,-max_abundance,-max_biomass,-shannon,-site_n,-season,-year)


one_observer<- one_observer %>% group_by(across(1:30)) %>% summarise("total_abundance"= sum(abundance)) %>% ungroup()

```

change columns names and values to fit the medata

```{r}

complexity_clean_data <-one_observer %>% rename("protection" = "reserve", 
                                                "site" = "location",
                                                "unique_trans_id" = "id",
                                                "sp.n" = "total_abundance",
                                                "sp.length" = "length")
complexity_clean_data$protection<-ifelse(complexity_clean_data$protection=="In","Protected","Not protected")

```

## save complexity data:

```{r}
#save(complexity_clean_data,file = "complexity_clean_data.Rdata")
```


