---
title: "5.1 dark diversity results"
author: "TG"
date: "2023-01-23"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

Abundance based Beals index

# Data prep and DD calculations

* load libraries

```{r}
library(tidyverse)
library(vegan)
library(plotrix)

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Dark diversity"
```

* load data

This is the **clean** version that have no NA in protection and no cryptic species

```{r}

load("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/clean_dd_medata.rdata")

clean_dd_medata<-medata # keep the original

```

convert to wide


```{r}
medata<-medata %>% select(1:22) # keep only metadate and abundances

medata<-medata %>% group_by(across(1:21)) %>% summarise(amount = sum(sp.n)) %>% ungroup() # sum ind from the same sp in each transect


```
low occupancy per basin

```{r}

medata <-medata %>% group_by(alt_basin,species) %>% mutate(occupancy = n())

medata<-medata %>% filter(occupancy > 5)

medata$occupancy<-NULL

```


```{r}
medata_wide<-spread(medata,species,amount,fill = 0) # convert to wide format

```

Convert the abundance data to probabilities using Beals index

note - we used presence absence data, type 0, include = False as done in previews papers

```{r}

beals <- medata_wide 

beals[,21:ncol(beals)] <- beals(beals[,21:ncol(beals)],type = 0,include = FALSE) 

```

gather the beal data and unite the species probability data with the abundance data 

```{r}
gather_beals <-gather(beals,"species","beal",21:ncol(beals))
gather_beals<-gather_beals %>% arrange(site,trans)

dark_diversity<-left_join(gather_beals,medata) 

dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species


```


Set the threshold for being included in the dark diversity:

the mean Beals probability for each **present** species in each basin in the data


```{r}

threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(beal_mean_prob_when_present = mean(beal),# mean beal value for the species when present
            beal_mean_plus_5 = beal_mean_prob_when_present + 0.05, # mean + 0.05 (old version)
            beal_quantile_5 =quantile(beal, probs = 0.05), # 5% quantile
            beal_quantile_1 = quantile(beal, probs = 0.01), # 1% quantile
            beal_quantile_10 =quantile(beal, probs = 0.1), # 10% quantile 
            beal_quantile_50 =quantile(beal, probs = 0.5), # 50 % quantile (median)
            min_beal = min(beal), # minimum value
            max_beal = max(beal), # maximum value
            n = n()) # number of transects used to set threshold


```

Test for one species in one basin


```{r}

threshold_test <- dark_diversity %>%
  filter(amount > 0,species =="Epinephelus.marginatus",alt_basin =="Levantine Sea") %>%
  mutate(beal_mean_prob_when_present = mean(beal),
            beal_quantile_5 =quantile(beal, probs = 0.05),
            beal_quantile_1 = quantile(beal, probs = 0.01),
            beal_quantile_10 =quantile(beal, probs = 0.1),
            beal_mean_plus_5 = beal_mean_prob_when_present +0.05,
            beals_median  = median(beal),
            min_beal = min(beal),
            max_beal = max(beal),
            n = n())



ggplot(data = threshold_test,aes(x= beal))+geom_histogram()+
geom_vline(xintercept = threshold_test$beal_quantile_5,col = "blue")+
geom_vline(xintercept = threshold_test$beal_mean_prob_when_present,col = "green")+
geom_vline(xintercept = threshold_test$beal_quantile_10,col = "brown")+
geom_vline(xintercept = threshold_test$beal_mean_plus_5,col = "red")
```


join with dark diversity data

```{r}

dark_diversity<- left_join(dark_diversity,threshold)

dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold


```


determine which species should be included in the dark diversity

`amount == 0` - the species is absent
`dark_diversity$beal > dark_diversity$beal_quantile_X` - its probability to occur is larger than X% quantile

```{r}

dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$beal_quantile_5,
                               "Dark diversity",
                               "Observed richness")

```


remove observation that are not on the site species pool, i.e., `status = Observed richness` and `amount = 0`

```{r}

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))


```

```{r}
ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")
```

remove species which threshold was determine by less than 5 occurrences 

```{r}
# dark_diversity <- dark_diversity %>% filter(n > 5)

# ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")

```


lets take a look at the dark assemblage

```{r}
only_dark <-dark_diversity %>% filter(species_status =="Dark diversity")

summary.factor(only_dark$protection)
summary(only_dark$enforcement)
summary.factor(only_dark$species)

```


# Dark diversity by protection

```{r}

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


ggplot(protection,aes(x = protection,y = richness, fill = type))+geom_boxplot()

protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()


protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()

```

# Dark diversity by enforcment

```{r}

enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


ggplot(enf,aes(x = enforcement,y = richness, fill = type))+geom_boxplot()

enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()



enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()

```
