---
title: "6.1 sensativity analyses"
author: "TG"
date: "2023-01-29"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Goals:

In this code I want to check my results sensativity to:

* tresholds: 5%,10% & 50% quantlies
* species occopancy: 3,5,10,20,50 (before and after calculating beals index)
* beals on abundances vs presnt/absance data
* separting israel from the levant basin


# Data prep and DD calculations

* load libraries

```{r}
library(tidyverse)
library(vegan)
library(plotrix)

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity"
plot_dir_st1<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_1"
```

* load data

This is the **clean** version that have no NA in protection and no cryptic species

```{r}

load("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/clean_dd_medata.rdata")

clean_dd_medata<-medata # keep the original

```

convert to wide and add column that separte israel


```{r}
medata<-medata %>% select(1:22) # keep only metadate and abundances

medata<-medata %>% group_by(across(1:21)) %>% summarise(amount = sum(sp.n)) %>% ungroup() # sum ind from the same sp in each transect


medata<-medata %>% mutate("alt_alt_basin"= alt_basin) %>% relocate(alt_alt_basin,.after = alt_basin)

medata$alt_alt_basin<-ifelse(medata$country == "Israel",
                             "Israel",
                             medata$alt_alt_basin)
```

# first stage

I will compare the diffrent occupany and quantile
while the beal will be set on present/abasence, israel will be included in the Levant and occupancy filtering will happened before beal calculation



```{r}

comper <- data.frame("occup"= rep(c(3,5,10,20,50),each = 3))
comper$quantile<-rep(c(5,10,50))
comper$quantile_prob<-rep(c(0.05,0.1,0.5))

comper$id<-paste0("oc",comper$occup,"_qu",comper$quantile)

```


```{r}

all_options<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_1\\",i)
  
  medata_loop <-medata %>% group_by(alt_basin,species) %>% mutate(occupancy = n())
  medata_loop<-medata_loop %>% filter(occupancy > oc)
  medata_loop$occupancy<-NULL
  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 0,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(quantile =quantile(beal, probs = prob), # 5% quantile
             n = n()) # number of transects used to set threshold

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

all_options[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}
```
# part 2

same but occurances are calculated after the beals

```{r}
all_options_oc_after_beal<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  oc
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_2\\",i)
  

  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 0,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(quantile =quantile(beal, probs = prob), # 5% quantile
             n = n()) # number of transects used to set threshold

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

dark_diversity <- dark_diversity %>% filter(n > oc)

all_options_oc_after_beal[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("b_af_","hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("b_af_obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("b_af_dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("b_af_obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("b_af_dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}

```

# part 3

now ill rerun previews loops for Beals based on abundance and not PA

occupancy is filterd before cakculating beals index
```{r}
pa_before_all_options<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_3\\before\\",i)
  
  medata_loop <-medata %>% group_by(alt_basin,species) %>% mutate(occupancy = n())
  medata_loop<-medata_loop %>% filter(occupancy > oc)
  medata_loop$occupancy<-NULL
  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 1,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(quantile =quantile(beal, probs = prob), # 5% quantile
             n = n()) # number of transects used to set threshold

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

pa_before_all_options[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("pa_before_hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("pa_before_obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("pa_before_dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("pa_before_obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("pa_before_dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}
```

occupancy is filterd after calculating beals index

```{r}
all_options_oc_after_beal_abundance<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_3\\after\\",i)
  

  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 1,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(quantile =quantile(beal, probs = prob), # 5% quantile
             n = n()) # number of transects used to set threshold

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

dark_diversity <- dark_diversity %>% filter(n > oc)

all_options_oc_after_beal_abundance[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("b_af_abun","hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("b_af_abun_obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("b_af_abun_dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("b_af_abun_obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("b_af_abun_dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}

```
# part 4

separate Israel...ill do that for occupancy filter before and after

## before

```{r}
israel_before<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_4\\before\\",i)
  
  medata_loop <-medata %>% group_by(alt_alt_basin,species) %>% mutate(occupancy = n())
  medata_loop<-medata_loop %>% filter(occupancy > oc)
  medata_loop$occupancy<-NULL
  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 0,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_alt_basin) %>% 
  summarise(quantile =quantile(beal, probs = prob),
             n = n()) 

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) 

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

israel_before[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("isr_before_hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("isr_before_obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("isr_before_dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("isr_before_obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("isr_before_dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}
```

## after

```{r}
isr_after<-list()

for (i in unique(comper$id)) {
  
  #set the conditions
  
  comp <-comper %>% filter(id == i)
  
  oc<-comp$occup
  prob<- comp$quantile_prob
  
  plot_dir_st1<-paste0("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\figures\\results\\Sensativity\\stage_4\\after\\",i)
  

  
  medata_wide<-spread(medata_loop,species,amount,fill = 0) # convert to wide format

  beals <- medata_wide 
  beals[,22:ncol(beals)] <- beals(beals[,22:ncol(beals)],type = 0,include = FALSE) 
 
  gather_beals <-gather(beals,"species","beal",22:ncol(beals))
  gather_beals<-gather_beals %>% arrange(site,trans)
  dark_diversity<-left_join(gather_beals,medata_loop) 
  dark_diversity$amount[is.na(dark_diversity$amount)]<-0 # now each transect have list of all species. add 0 instead of NA to the newly created species

  threshold <- dark_diversity %>%
  filter(amount >0) %>% 
  group_by(species,alt_alt_basin) %>% # set threshold for each species & basin (alt_basin) 
  summarise(quantile =quantile(beal, probs = prob), # 5% quantile
             n = n()) # number of transects used to set threshold

  dark_diversity<- left_join(dark_diversity,threshold)

  dark_diversity<-dark_diversity %>% drop_na(n) # remove the species that did not occur at all in a given basin and thus don't have a threshold

  
dark_diversity$species_status <- ifelse(dark_diversity$amount == "0" & dark_diversity$beal > dark_diversity$quantile,
                               "Dark diversity",
                               "Observed richness")

dark_diversity<- dark_diversity %>% filter(!(species_status == "Observed richness" & amount == 0))

dark_diversity$occup<-rep(oc)
dark_diversity$quantile_prob<-rep(prob)

dark_diversity <- dark_diversity %>% filter(n > oc)

isr_after[[i]]<-dark_diversity

#plots:

hist <- ggplot(data = dark_diversity,aes(x= n,fill = species_status))+geom_histogram(alpha = 0.5,color = "black")+ggtitle(i)

ggsave(hist,filename = paste0("isr_af_abun","hist_",i,".png"),path = plot_dir_st1)


# plot by protection

protection <- dark_diversity %>%
  group_by(unique_trans_id,protection,species_status) %>% 
  summarise(richness = n_distinct(species)) %>% 
  arrange(unique_trans_id)

  
protection<-spread(protection,species_status,richness,fill = 0)

protection$total<-protection$`Dark diversity` + protection$`Observed richness`


protection <- gather(protection,type,richness,3:5)
protection <- protection %>% arrange(unique_trans_id)


protection <- protection %>%
  group_by(protection,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 


protection %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=protection,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+
  ggtitle(i)+
  theme(legend.position="none")

ggsave(filename = paste("isr_af_obs_dd_protection_",i,".png"),path = plot_dir_st1)

protection %>%  filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=protection,y = mean_richness)+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean DD richness per transect")+
  theme_classic()+ggtitle(i)

 ggsave(filename = paste("isr_af_dd_protection_",i,".png"),path = plot_dir_st1)
 
 
 enf<-dark_diversity %>%
  group_by(enforcement,species_status,unique_trans_id) %>%
  summarise(richness = n_distinct(species)) %>% arrange(unique_trans_id)


enf<-spread(enf,species_status,richness,fill = 0)

enf$total<-enf$`Dark diversity` + enf$`Observed richness`

enf <- gather(enf,type,richness,3:5)
enf <- enf %>% arrange(unique_trans_id)


enf <- enf %>%
  group_by(enforcement,type) %>%
  summarise("mean_richness"= mean(richness),
            "se"=std.error(richness)) 

enf %>%  na.omit() %>% filter(type != "total") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity")+ylab("Mean richness per transect")+
  scale_fill_manual(values = c("#1B2D48","#F7D54A"))+theme_classic()+ggtitle(i)+
  theme(legend.position="none")

 ggsave(filename = paste("isr_af_obs_dd_enforcment_",i,".png"),path = plot_dir_st1)


enf %>%  na.omit() %>% filter(type == "Dark diversity") %>% 
ggplot()+
       aes(x=enforcement,y = mean_richness,fill = type )+
  geom_bar(stat="identity",position=position_dodge(.9),fill = "#1B2D48")+
  geom_errorbar(aes(ymin=mean_richness-se, ymax=mean_richness+se), width=.2,
                 position=position_dodge(.9))+ylab("Mean Dark diversity per transect") +theme_classic()+ggtitle(i)

 ggsave(filename = paste("isr_af_dd_enforcment_",i,".png"),path = plot_dir_st1)

print(i)
}

```

