---
title: "9.1 log ratio plot"
author: "TG"
date: "2023-06-11"
output: html_document
---


```{r}
library(DarkDiv)
library(tidyverse)
library(vegan)
library(plotrix)
library(sjPlot)
library(sjmisc)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(SingleCaseES)
select <- dplyr::select

load("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/clean_dd_medata.rdata")

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"
```

## Final data prepartion

```{r}

clean_dd_medata<-medata # keep the original

medata<-medata %>% select(1:23) # keep only metadata and abundances

medata<-medata %>% group_by(across(1:22)) %>% summarise(amount = sum(sp.n)) %>% ungroup() # sum ind from the same sp in each transect

medata$alt_basin<- ifelse(medata$country == "Israel","Israel",medata$alt_basin) # Sep Israel

medata<-medata %>% filter(country != "Cyprus") # Remove cyprus data


medata$alt_basin[medata$new_site_name=="cabrera"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin[medata$new_site_name=="illa_de_l'aire"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin<-ifelse(medata$alt_basin == "Levantine Sea","Mediterranean Sea - Eastern Basin",medata$alt_basin)# restore turkey transect to previos basin

medata_wide<-spread(medata,species,amount,fill = 0) # convert to wide format

# remove this site - outlyar since there is feeding of groupers in the site

medata_wide<-medata_wide %>% filter(new_site_name !="Crete_Kalydon")

```



```{r}
index_list<-list()
data_list<-list()

for (i in unique(medata_wide$alt_basin)) {
 
   basin_data<-medata_wide %>% filter(alt_basin == i)
   
   sp_matrix<-basin_data[,22:ncol(basin_data)]
   
   sp_matrix<-sp_matrix[, colSums(sp_matrix != 0) > 0] # Remove species with all 0
  
   dd<-DarkDiv(sp_matrix,method = "Hypergeometric")
   
   index_list[[i]]<-dd
   
   dd_dark<-dd[["Dark"]]
   dd_dark<-as.data.frame(dd_dark)
   
   dd_dark[dd_dark<0.9]<-0 # Set threshold
   dd_dark[dd_dark>0]<-1 # convert to 1 species above the th
   
   dd_dark<-dd_dark %>% mutate("dark"=rowSums(dd_dark,na.rm = T))
   dd_dark_sum<- dd_dark %>% select(dark)
   
   dd_pool<-dd[["Pool"]]
   dd_pool<-as.data.frame(dd_pool)
  
   dd_pool[dd_pool<0.9]<-0
   dd_pool[dd_pool>0]<-1
   
   dd_pool<-dd_pool %>% mutate("total"=rowSums(dd_pool,na.rm = T))
   dd_pool_sum<-dd_pool %>% select(total)
   
   medata_dd<-basin_data[,1:21]
   medata_dd<-as.data.frame(medata_dd)

   medata_dd<-bind_cols(medata_dd,dd_pool_sum)
   medata_dd<-bind_cols(medata_dd,dd_dark_sum)

   medata_dd$observed_richness<-medata_dd$total-medata_dd$dark
   
   data_list[[i]]<-medata_dd
}

dd_data<-bind_rows(data_list)
```

clean 

```{r}
rm(list=setdiff(ls(), "dd_data"))

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"

labl<-c(dark = "Dark diversity",
        observed_richness = "Observed richness",
        total="Total species pool")

pair_list<-list()
site_list<-list()
```



* brijoni


```{r}
Brijoni <- dd_data %>% filter(site == "brijuni")

Brijoni<-gather(Brijoni,"index","richness",22:24)

Brijoni$pair<-"Brijoni"

site_list[[1]]<-Brijoni

Brijoni <- Brijoni %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Brijoni$pair<-"Brijoni"

pair_list[[1]]<-Brijoni
```

* israel

```{r}
isr<-dd_data %>% filter(country == "Israel",site != "habonim")

isr<-gather(isr,"index","richness",22:24)
isr$pair<-isr$site
site_list[[2]]<-isr

isr <- isr %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

isr$pair<-isr$site

pair_list[[2]]<-isr

```

* France

```{r}
Banyuls <- dd_data %>% filter(new_site_name == "Banyuls")

Banyuls<-gather(Banyuls,"index","richness",22:24)

Banyuls$pair <-"Banyuls"

site_list[[3]]<-Banyuls

Banyuls <- Banyuls %>% group_by(country,new_site_name,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Banyuls$pair <-"Banyuls"

pair_list[[3]]<-Banyuls
```

* spain

note - there are two possible control sites for Cabrera MPA, I used Dragonera because it a bit closer and also an island  and omit formentor *but* we might change it later or use 2 pairs (or aggregate the two control sites...)

note 2- same goes for cap_de_creus - no pair so i ommited but we should talk to the team and see what they think.

```{r}
spain<-dd_data %>% filter(country == "Spain")

spain<-gather(spain,"index","richness",22:24)

spain <- spain %>% mutate(pair = case_when(
  new_site_name == "formentera"~ "Formentera",
  new_site_name == "eivissa"~ "Formentera",
  new_site_name == "cabrera"~"Cabrera",
  new_site_name == "dragonera"~ "Cabrera",
  new_site_name == "cavalleria"~ "Cavalleria",
  new_site_name == "illa_de_l'aire"~ "Cavalleria",
  new_site_name == "medes"~ "Medes"
  
  ))

spain<-spain %>% drop_na(pair)


site_list[[4]]<-spain

spain$site<-spain$pair

spain <- spain %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

spain$pair <- spain$site

pair_list[[4]] <- spain

```

Italy

Need to decide how to treat those 3 sites...

  new_site_name == "carloforte"~"pair_2" , # control
  new_site_name == "asinara"~ "pair_2", # protected 3 & 2
  new_site_name == "tavolara"~ "pair_2", # protected 3
  
i did two different comparison with carloforta as the control for both MPAs  
  
  

```{r}
itl<-dd_data %>% filter(country == "Italy",data.origin != "Belmaker")


itl<-gather(itl,"index","richness",22:24)

itl <- itl %>% mutate(pair = case_when(
  new_site_name == "portofino"~ "Portofino",
  new_site_name == "genova"~ "Portofino",
  new_site_name == "torre_guaceto"~ "Torre guaceto",
  new_site_name == "otranto"~ "Torre guaceto",
  new_site_name == "asinara"~ "Asinara", 
  new_site_name == "carloforte"~"Asinara"
  ))

itl<-itl %>% drop_na(pair)

itl_2<-dd_data %>% filter(country == "Italy",new_site_name %in% c("tavolara","carloforte"))

itl_2<-gather(itl_2,"index","richness",22:24)

itl_2$pair = "Tavolara"

itl<-rbind(itl,itl_2)

site_list[[5]]<-itl
itl$site<-itl$pair

itl <- itl %>% group_by(country,site,protection,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

itl$pair <- itl$site

pair_list[[5]] <- itl


```


```{r}
pairs<-bind_rows(pair_list)


log_ratio<-pairs %>% ungroup() %>%
  filter(index == "dark") %>% select(country,pair,index,protection,mean_richness) %>% 
  spread(protection,mean_richness) %>% 
  mutate("ratio"= `Protected`/`Not protected`) %>% 
  mutate("log_ratio" = log(ratio)) %>% 
  mutate(pair = str_to_title(pair))

log_ratio$pair<-as.factor(log_ratio$pair)

log_ratio<-log_ratio%>% mutate(pair = fct_reorder(pair,log_ratio))

ggplot(data = log_ratio,aes(x=log_ratio,y = pair))+geom_point(color = '#7DCDE4',size = 4)+theme_classic()+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  ylab("MPA")+xlab("Log ratio")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())


```



# CalcSpeciesSE log ratios

```{r}

log_data_raw<-bind_rows(site_list)

log_data<-log_data_raw %>% select(pair,protection,index,richness) %>% filter(index != "total")

ratio = batch_calc_ES(dat = log_data, grouping =  c(pair,index), condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

ratio$pair = str_to_title(ratio$pair)
ratio$pair<-as.factor(ratio$pair)

ratio<-ratio%>% mutate(pair = fct_reorder(pair,ave(index =="dark",Est),max))

levels(d_ratio$pair)

ratio<-ratio%>% mutate(pair = fct_reorder(pair,d_ratio$Est))

ggplot(data = ratio,aes(x=Est,y = pair,color = index))+theme_classic()+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  ylab("MPA")+xlab("Log ratio")+geom_errorbar(aes(xmin= CI_lower, xmax=CI_upper), width=.2)+geom_point(size = 4)+scale_color_manual(values=c('#7DCDE4',"#F2E96C"))+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())



ggplot(data = ratio,aes(x=Est,y = pair))+theme_classic()+
  ylab("MPA")+xlab("Log ratio")+geom_errorbar(aes(xmin= Est-SE, xmax=Est+SE), width=.2)+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+geom_point(size = 4)+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())


#ggsave("ratio_se.png",path = plot_dir,width = 12,height = 7)
#ggsave("ratio_se.pdf",path = plot_dir,width = 12,height = 7)

```


```{r}

log_data_raw<-bind_rows(site_list)

log_data<-log_data_raw %>% select(pair,protection,index,richness) %>% filter(index != "total")

log_data$pair<-as.factor(log_data$pair)

log_data$pair<-str_to_title(log_data$pair)

# dark

d_ratio<-log_data %>% filter(index == "dark")

d_ratio = batch_calc_ES(dat = d_ratio, grouping =  pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase	 = "Not protected")

d_ratio<-d_ratio%>% mutate(pair = fct_reorder(pair,Est))

d_ratio$Index = "Dark diversity"
#observed

o_ratio<-log_data %>% filter(index == "observed_richness")

o_ratio = batch_calc_ES(dat = o_ratio, grouping = pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

o_ratio$Index<-"Observed richness"

#plot

ggplot(data = d_ratio,aes(x=Est,y=pair,color = Index))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 4)+
  geom_point(data = o_ratio,aes(x=Est,y= pair,color=Index),size = 4)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+
  geom_errorbar(data = d_ratio,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = o_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C"))+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())

#ggsave("ratio_se_dark_obs.png",path = plot_dir,width = 12,height = 7)
#ggsave("ratio_se_dark_obs.pdf",path = plot_dir,width = 12,height = 7)

ggplot(data = d_ratio,aes(x=Est,y=pair,color = Index))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 3)+
  geom_point(data = o_ratio,aes(x=Est,y= pair,color=Index),size = 3)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+
  geom_errorbar(data = d_ratio,aes(xmin= CI_lower, xmax=CI_upper), width=.2)+
  geom_errorbar(data = o_ratio ,aes(xmin= CI_lower, xmax=CI_upper), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C"))+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())

```

