---
title: "9.3 log ratio with metafor summary"
author: "TG"
date: "2023-09-05"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
library(DarkDiv)
library(tidyverse)
library(vegan)
library(plotrix)
library(sjPlot)
library(sjmisc)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(SingleCaseES)
library(metafor)
select <- dplyr::select

load("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/clean_dd_medata.rdata")

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"
```

## Final data prepartion

```{r}

clean_dd_medata<-medata # keep the original

medata<-medata %>% select(1:23) # keep only metadata and abundances

medata<-medata %>% group_by(across(1:22)) %>% summarise(amount = sum(sp.n)) %>% ungroup() # sum ind from the same sp in each transect

medata$alt_basin<- ifelse(medata$country == "Israel","Israel",medata$alt_basin) # Sep Israel

medata<-medata %>% filter(country != "Cyprus") # Remove cyprus data


medata$alt_basin[medata$new_site_name=="cabrera"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin[medata$new_site_name=="illa_de_l'aire"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin<-ifelse(medata$alt_basin == "Levantine Sea","Mediterranean Sea - Eastern Basin",medata$alt_basin)# restore turkey transect to previos basin

medata_wide<-spread(medata,species,amount,fill = 0) # convert to wide format

# remove this site - outlyar since there is feeding of groupers in the site

medata_wide<-medata_wide %>% filter(new_site_name !="Crete_Kalydon")

```



```{r}
index_list<-list()
data_list<-list()

for (i in unique(medata_wide$alt_basin)) {
 
   basin_data<-medata_wide %>% filter(alt_basin == i)
   
   sp_matrix<-basin_data[,22:ncol(basin_data)]
   
   sp_matrix<-sp_matrix[, colSums(sp_matrix != 0) > 0] # Remove species with all 0
  
   dd<-DarkDiv(sp_matrix,method = "Hypergeometric")
   
   index_list[[i]]<-dd
   
   dd_dark<-dd[["Dark"]]
   dd_dark<-as.data.frame(dd_dark)
   
   dd_dark[dd_dark<0.9]<-0 # Set threshold
   dd_dark[dd_dark>0]<-1 # convert to 1 species above the th
   
   dd_dark<-dd_dark %>% mutate("dark"=rowSums(dd_dark,na.rm = T))
   dd_dark_sum<- dd_dark %>% select(dark)
   
   dd_pool<-dd[["Pool"]]
   dd_pool<-as.data.frame(dd_pool)
  
   dd_pool[dd_pool<0.9]<-0
   dd_pool[dd_pool>0]<-1
   
   dd_pool<-dd_pool %>% mutate("total"=rowSums(dd_pool,na.rm = T))
   dd_pool_sum<-dd_pool %>% select(total)
   
   medata_dd<-basin_data[,1:21]
   medata_dd<-as.data.frame(medata_dd)

   medata_dd<-bind_cols(medata_dd,dd_pool_sum)
   medata_dd<-bind_cols(medata_dd,dd_dark_sum)

   medata_dd$observed_richness<-medata_dd$total-medata_dd$dark
   
   data_list[[i]]<-medata_dd
}

dd_data<-bind_rows(data_list)
```


make a table for the ms


```{r}

table_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\tables"


trans_table<-dd_data %>% group_by(country,new_site_name,protection) %>% summarise("n"= n(), "Mean observed richness"= round(mean(observed_richness),1),"Mean dark diversity" = round(mean(dark),1))

trans_table<- trans_table %>% arrange(country,new_site_name,protection)

colnames(trans_table)<- c("Country","Site","Protection","Number of transects","Mean observed richness","Mean dark diversity")

trans_table$Site <- sub("_", " ", trans_table$Site)
trans_table$Site<-str_to_title(trans_table$Site)

#write.csv(trans_table,paste0(table_dir,"\\transect_per_site.csv"))
```





clean 

```{r}
rm(list=setdiff(ls(), "dd_data"))

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"

labl<-c(dark = "Dark diversity",
        observed_richness = "Observed richness",
        total="Total species pool")

pair_list<-list()
site_list<-list()
```



* brijoni


```{r}
Brijoni <- dd_data %>% filter(site == "brijuni")

Brijoni<-gather(Brijoni,"index","richness",22:24)

Brijoni$pair<-"Brijoni"

site_list[[1]]<-Brijoni

Brijoni <- Brijoni %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Brijoni$pair<-"Brijoni"

pair_list[[1]]<-Brijoni
```

* israel

```{r}
isr<-dd_data %>% filter(country == "Israel",site != "habonim")

isr<-gather(isr,"index","richness",22:24)
isr$pair<-isr$site
site_list[[2]]<-isr

isr <- isr %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

isr$pair<-isr$site

pair_list[[2]]<-isr

```

* France

```{r}
Banyuls <- dd_data %>% filter(new_site_name == "Banyuls")

Banyuls<-gather(Banyuls,"index","richness",22:24)

Banyuls$pair <-"Banyuls"

site_list[[3]]<-Banyuls

Banyuls <- Banyuls %>% group_by(country,new_site_name,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Banyuls$pair <-"Banyuls"

pair_list[[3]]<-Banyuls
```

* spain

note - there are two possible control sites for Cabrera MPA, I used Dragonera because it a bit closer and also an island  and omit formentor *but* we might change it later or use 2 pairs (or aggregate the two control sites...)

note 2- same goes for cap_de_creus - no pair so i ommited but we should talk to the team and see what they think.

```{r}
spain<-dd_data %>% filter(country == "Spain")

spain<-gather(spain,"index","richness",22:24)

spain <- spain %>% mutate(pair = case_when(
  new_site_name == "formentera"~ "Formentera",
  new_site_name == "eivissa"~ "Formentera",
  new_site_name == "cabrera"~"Cabrera",
  new_site_name == "dragonera"~ "Cabrera",
  new_site_name == "cavalleria"~ "Cavalleria",
  new_site_name == "illa_de_l'aire"~ "Cavalleria",
  new_site_name == "medes"~ "Medes"
  
  ))

spain<-spain %>% drop_na(pair)


site_list[[4]]<-spain

spain$site<-spain$pair

spain <- spain %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

spain$pair <- spain$site

pair_list[[4]] <- spain

```

Italy

Need to decide how to treat those 3 sites...

  new_site_name == "carloforte"~"pair_2" , # control
  new_site_name == "asinara"~ "pair_2", # protected 3 & 2
  new_site_name == "tavolara"~ "pair_2", # protected 3
  
i did two different comparison with carloforta as the control for both MPAs  
  
  

```{r}
itl<-dd_data %>% filter(country == "Italy",data.origin != "Belmaker")


itl<-gather(itl,"index","richness",22:24)

itl <- itl %>% mutate(pair = case_when(
  new_site_name == "portofino"~ "Portofino",
  new_site_name == "genova"~ "Portofino",
  new_site_name == "torre_guaceto"~ "Torre guaceto",
  new_site_name == "otranto"~ "Torre guaceto",
  new_site_name == "asinara"~ "Asinara", 
  new_site_name == "carloforte"~"Asinara"
  ))

itl<-itl %>% drop_na(pair)

itl_2<-dd_data %>% filter(country == "Italy",new_site_name %in% c("tavolara","carloforte"))

itl_2<-gather(itl_2,"index","richness",22:24)

itl_2$pair = "Tavolara"

itl<-rbind(itl,itl_2)

site_list[[5]]<-itl
itl$site<-itl$pair

itl <- itl %>% group_by(country,site,protection,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

itl$pair <- itl$site

pair_list[[5]] <- itl


```



```{r}
log_data_raw<-bind_rows(site_list)

rm(list=setdiff(ls(), "log_data_raw"))

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"

labl<-c(dark = "Dark diversity",
        observed_richness = "Observed richness",
        total="Total species pool")

```

```{r}

log_data<-log_data_raw %>% select(pair,protection,index,richness) 

log_data$pair<-as.factor(log_data$pair)

log_data$pair<-str_to_title(log_data$pair)

# dark

d_ratio<-log_data %>% filter(index == "dark")

d_ratio = batch_calc_ES(dat = d_ratio, grouping =  pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase	 = "Not protected")

d_ratio<-d_ratio%>% mutate(pair = fct_reorder(pair,Est))

d_ratio$Index = "Dark diversity"

#observed

o_ratio<-log_data %>% filter(index == "observed_richness")

o_ratio = batch_calc_ES(dat = o_ratio, grouping = pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

o_ratio$Index<-"Observed richness"

# Total

t_ratio<-log_data %>% filter(index == "total")

t_ratio = batch_calc_ES(dat = t_ratio, grouping = pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

t_ratio$Index<-"Total richness"

#plot

ggplot(data = d_ratio,aes(x=Est,y=pair,color = Index))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 4)+
  geom_point(data = o_ratio,aes(x=Est,y= pair,color=Index),size = 4)+
  geom_point(data = t_ratio,aes(x=Est,y= pair,color=Index),size = 4)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+
  geom_errorbar(data = d_ratio,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = o_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = t_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C","#ACD672"))+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20),
        strip.background = element_blank())

#ggsave("ratio_se_dark_obs.png",path = plot_dir,width = 12,height = 7)
#ggsave("ratio_se_dark_obs.pdf",path = plot_dir,width = 12,height = 7)

```

```{r}
protection<-log_data_raw %>% select(pair,enforcement) %>% distinct(,.keep_all = T) %>% filter(as.numeric(as.character(enforcement))>2)


```


add meta-analyses summery from metafor


```{r}
metafor_data<-log_data_raw

metafor_data$protection<-ifelse(metafor_data$protection=="Protected","MPA","control")

```


```{r}

metafor_dark<-metafor_data %>%
  filter(index =="dark") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 

RR_effect_sizes_dark <- escalc( # Function for calculating effect sizes.
    "ROM",
    # Specify the effect size we want to calculate. In this case ROM for the Ratio of Means or Response Ratio
    m1i = mean_MPA,              # MPA mean richness  
    n1i = n_MPA,                 # MPA sample size
    sd1i = sd_MPA,               # MPA SD
    m2i = mean_control,          # control mean richness
    n2i = n_control,             # control sample size
    sd2i = sd_control,           # control site SD
    var.names=c("yi","vi","sei"),
    data = metafor_dark ,    # This is where the escalc function can find all the data for our meta-analysis
  )


RR_effect_sizes_dark$se = sqrt(RR_effect_sizes_dark$vi)


random_effect_model_results <- rma(yi, # this is the effect size from each row in database
                                   vi, # measure of variance from each row in database
                                   method = "REML", # Specifies fixed effects model
                                   slab = pair, #  prepares  labels for the forest plot 
                                   data = RR_effect_sizes_dark) # Let R know the dataframe we'll use for our model calculations

random_effect_model_results


very_protected_dark <-RR_effect_sizes_dark %>% filter(pair %in% protection$pair) 


pro_random_effect_model_results <- rma(yi, 
                                   vi, 
                                   method = "REML", 
                                   slab = pair, 
                                   data = very_protected_dark)


pro_random_effect_model_results

forest(
  RR_effect_sizes_dark$yi, # These are effect sizes from each row in database
  RR_effect_sizes_dark$vi, # These are variances from each row in database
  annotate = T,            # Setting this to false prevents R from including CIs for each effect sizes 
  order="obs",             # set the order of the sites 
  slab = random_effect_model_results$slab, # A individual effect size labeled 
  xlab = "ln(Response Ratio)", # Label for x-axis
  cex = .8,                    # Text side for study labels
  pch = 15,                    # shape of bars in forest plot
  cex.lab = 1                  # Size of x-axis label
)

```

```{r}

d_ratio<-d_ratio %>% arrange(desc(Est)) %>% mutate(order = seq(1,nrow(d_ratio),1))



d_summary<-d_ratio[1,]

d_summary$pair <- "Summary"

d_summary$SE<-random_effect_model_results$se

d_summary$Est<-random_effect_model_results$b[[1]]

d_summary$CI_lower<-NA
d_summary$CI_upper<-NA
d_summary$order<-15

d_ratio<-rbind(d_ratio,d_summary)


d_ratio<-d_ratio%>% mutate(pair = fct_reorder( pair,desc(order)))


d_ratio$shape<-ifelse(d_ratio$pair == "Summary","Sum","Regular")

d_ratio$shape<-as.factor(d_ratio$shape)
```


```{r}
metafor_obs<-metafor_data %>%
  filter(index =="observed_richness") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 



RR_effect_sizes_obs <- escalc( # Function for calculating effect sizes.
    "ROM",
    # Specify the effect size we want to calculate. In this case ROM for the Ratio of Means or Response Ratio
    m1i = mean_MPA,              # MPA mean richness  
    n1i = n_MPA,                 # MPA sample size
    sd1i = sd_MPA,               # MPA SD
    m2i = mean_control,          # control mean richness
    n2i = n_control,             # control sample size
    sd2i = sd_control,           # control site SD
    data = metafor_obs # This is where the escalc function can find all the data for our meta-analysis
  )



obs_random_effect_model_results <- rma(yi, # this is the effect size from each row in database
                                   vi, # measure of variance from each row in database
                                   method = "REML", # Specifies fixed effects model
                                   slab = pair, #  prepares  labels for the forest plot 
                                   data = RR_effect_sizes_obs) # Let R know the dataframe we'll use for our model calculations

obs_random_effect_model_results

very_protected_obs <-RR_effect_sizes_obs %>% filter(pair %in% protection$pair) 


pro_obs_random_effect_model_results <- rma(yi, 
                                   vi, 
                                   method = "REML", 
                                   slab = pair, 
                                   data = very_protected_obs)

pro_obs_random_effect_model_results



o_summary<-o_ratio[1,]

o_summary$pair <- "Summary"

o_summary$SE<-obs_random_effect_model_results$se

o_summary$Est<-obs_random_effect_model_results$b[[1]]

o_summary$CI_lower<-NA
o_summary$CI_upper<-NA

o_ratio<-rbind(o_ratio,o_summary)

o_ratio$shape<-ifelse(o_ratio$pair == "Summary","Sum","Regular")

o_ratio$shape<-as.factor(o_ratio$shape)
```





```{r}
metafor_total<-metafor_data %>%
  filter(index =="total") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 



RR_effect_sizes_total <- escalc( # Function for calculating effect sizes.
    "ROM",
    # Specify the effect size we want to calculate. In this case ROM for the Ratio of Means or Response Ratio
    m1i = mean_MPA,              # MPA mean richness  
    n1i = n_MPA,                 # MPA sample size
    sd1i = sd_MPA,               # MPA SD
    m2i = mean_control,          # control mean richness
    n2i = n_control,             # control sample size
    sd2i = sd_control,           # control site SD
    data = metafor_total # This is where the escalc function can find all the data for our meta-analysis
  )



total_random_effect_model_results <- rma(yi, # this is the effect size from each row in database
                                   vi, # measure of variance from each row in database
                                   method = "REML", # Specifies fixed effects model
                                   slab = pair, #  prepares  labels for the forest plot 
                                   data = RR_effect_sizes_total) # Let R know the dataframe we'll use for our model calculations

total_random_effect_model_results

very_protected_total <-RR_effect_sizes_total %>% filter(pair %in% protection$pair) 


pro_total_random_effect_model_results <- rma(yi, 
                                   vi, 
                                   method = "REML", 
                                   slab = pair, 
                                   data = very_protected_total)

pro_total_random_effect_model_results

t_summary<-t_ratio[1,]

t_summary$pair <- "Summary"

t_summary$SE<-total_random_effect_model_results$se

t_summary$Est<-total_random_effect_model_results$b[[1]]

t_summary$CI_lower<-NA
t_summary$CI_upper<-NA

t_ratio<-rbind(t_ratio,t_summary)

t_ratio$shape<-ifelse(t_ratio$pair == "Summary","Sum","Regular")

t_ratio$shape<-as.factor(t_ratio$shape)
```


```{r}
ggplot(data = d_ratio,aes(x=Est,y=pair,color = Index,shape = shape))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 6)+
  geom_point(data = o_ratio,aes(x=Est,y= pair,color=Index,shape = shape),size = 6)+
  geom_point(data = t_ratio,aes(x=Est,y= pair,color=Index,shape = shape),size = 6)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+
  geom_errorbar(data = d_ratio,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = o_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = t_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C","#ACD672"))+
  theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())+
  scale_shape_manual(values=c(16, 15))
  

#ggsave("ratio_se_dark_obs_total.png",path = plot_dir,width = 12,height = 8)
#ggsave("ratio_se_dark_obs_total.pdf",path = plot_dir,width = 12,height = 8)

```


# completeness index

```{r}
completness_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\completness"
```


```{r}
completeness<-log_data %>% group_by(pair,protection,index) %>% summarise("mean_richness"= mean(richness))


completeness<-spread(completeness,index,mean_richness)

completeness<-completeness %>% mutate("completness" = log(observed_richness/dark))


completeness_ratio<- completeness %>% select(pair,protection,completness) 

completeness_ratio<- spread(completeness_ratio,protection,completness)

 completeness_ratio<-completeness_ratio%>% 
  mutate("ratio"= `Protected`/`Not protected`) %>% 
  mutate("log_ratio" = log(ratio),"gap" = `Protected`-`Not protected`)

completeness_ratio<-completeness_ratio %>% mutate(pair = fct_reorder(pair,gap))

ggplot(completeness_ratio,aes(x=gap,y=fct_reorder(pair,gap)))+geom_point(size =5,color = "darkred")+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  ylab("MPA")+
  xlab("Completeness gap")+
  theme_classic()+theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())

#ggsave("completness gap.png",path = completness_dir,width = 12,height = 8)
#ggsave("completness gap.pdf",path = completness_dir,width = 12,height = 8)

```

```{r}

completeness<-log_data %>% group_by(pair,protection,index) %>% summarise("mean_richness"= mean(richness),n=n())


completeness<-spread(completeness,index,mean_richness)

completeness<-completeness %>% mutate("com" = log(observed_richness/dark))


completeness_met<- completeness %>% select(pair,protection,com,n) 

completeness_met<- completeness_met %>% pivot_wider(names_from=protection, 
              values_from=c(com, 
                            n)) 


com_effect <- escalc( # Function for calculating effect sizes.
    "MD",             # Specify the effect size we want to calculate.
    m1i = com_Protected,              # MPA mean richness  
    n1i = n_Protected,                 # MPA sample size
    m2i = `com_Not protected`,          # control mean richness
    n2i = `n_Not protected`,             # control sample size
    data = completeness_met #  data for our meta-analysis
  )



```

