---
title: "9.4 log ratio with metafor summary"
author: "TG"
date: "2023-09-26"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
editor_options: 
  chunk_output_type: console
---

In this script I orgenize the code from 9.3 to create the new log plot


```{r}
library(DarkDiv)
library(tidyverse)
library(vegan)
library(plotrix)
library(sjPlot)
library(sjmisc)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(SingleCaseES)
library(metafor)
select <- dplyr::select

load("G:/My Drive/PHD/chapter 3 MPA dd/Meddata-DD/clean_dd_medata.rdata")

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"
```

## Final data prepartion

```{r}

clean_dd_medata<-medata # keep the original

medata<-medata %>% select(1:23) # keep only metadata and abundances

medata<-medata %>% group_by(across(1:22)) %>% summarise(amount = sum(sp.n)) %>% ungroup() # sum ind from the same sp in each transect

medata$alt_basin<- ifelse(medata$country == "Israel","Israel",medata$alt_basin) # Sep Israel

medata<-medata %>% filter(country != "Cyprus") # Remove cyprus data


medata$alt_basin[medata$new_site_name=="cabrera"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin[medata$new_site_name=="illa_de_l'aire"]<-"Balearic (Iberian Sea)" # one basin for the site

medata$alt_basin<-ifelse(medata$alt_basin == "Levantine Sea","Mediterranean Sea - Eastern Basin",medata$alt_basin)# restore turkey transect to previos basin

medata_wide<-spread(medata,species,amount,fill = 0) # convert to wide format

# remove this site - outlyar since there is feeding of groupers in the site

medata_wide<-medata_wide %>% filter(new_site_name !="Crete_Kalydon")

```



```{r}
index_list<-list()
data_list<-list()

for (i in unique(medata_wide$alt_basin)) {
 
   basin_data<-medata_wide %>% filter(alt_basin == i)
   
   sp_matrix<-basin_data[,22:ncol(basin_data)]
   
   sp_matrix<-sp_matrix[, colSums(sp_matrix != 0) > 0] # Remove species with all 0
  
   dd<-DarkDiv(sp_matrix,method = "Hypergeometric")
   
   index_list[[i]]<-dd
   
   dd_dark<-dd[["Dark"]]
   dd_dark<-as.data.frame(dd_dark)
   
   dd_dark[dd_dark<0.9]<-0 # Set threshold
   dd_dark[dd_dark>0]<-1 # convert to 1 species above the th
   
   dd_dark<-dd_dark %>% mutate("dark"=rowSums(dd_dark,na.rm = T))
   dd_dark_sum<- dd_dark %>% select(dark)
   
   dd_pool<-dd[["Pool"]]
   dd_pool<-as.data.frame(dd_pool)
  
   dd_pool[dd_pool<0.9]<-0
   dd_pool[dd_pool>0]<-1
   
   dd_pool<-dd_pool %>% mutate("total"=rowSums(dd_pool,na.rm = T))
   dd_pool_sum<-dd_pool %>% select(total)
   
   medata_dd<-basin_data[,1:21]
   medata_dd<-as.data.frame(medata_dd)

   medata_dd<-bind_cols(medata_dd,dd_pool_sum)
   medata_dd<-bind_cols(medata_dd,dd_dark_sum)

   medata_dd$observed_richness<-medata_dd$total-medata_dd$dark
   
   data_list[[i]]<-medata_dd
}

dd_data<-bind_rows(data_list)
```

clean 

```{r}
rm(list=setdiff(ls(), "dd_data"))

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"

labl<-c(dark = "Dark diversity",
        observed_richness = "Observed richness",
        total="Total species pool")

pair_list<-list()
site_list<-list()
```



* brijoni


```{r}
Brijoni <- dd_data %>% filter(site == "brijuni")

Brijoni<-gather(Brijoni,"index","richness",22:24)

Brijoni$pair<-"Brijoni"

site_list[[1]]<-Brijoni

Brijoni <- Brijoni %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Brijoni$pair<-"Brijoni"

pair_list[[1]]<-Brijoni
```

* israel

```{r}
isr<-dd_data %>% filter(country == "Israel",site != "habonim")

isr<-gather(isr,"index","richness",22:24)
isr$pair<-isr$site
site_list[[2]]<-isr

isr <- isr %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

isr$pair<-isr$site

pair_list[[2]]<-isr

```

* France

```{r}
Banyuls <- dd_data %>% filter(new_site_name == "Banyuls")

Banyuls<-gather(Banyuls,"index","richness",22:24)

Banyuls$pair <-"Banyuls"

site_list[[3]]<-Banyuls

Banyuls <- Banyuls %>% group_by(country,new_site_name,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

Banyuls$pair <-"Banyuls"

pair_list[[3]]<-Banyuls
```

* spain

note - there are two possible control sites for Cabrera MPA, I used Dragonera because it a bit closer and also an island  and omit formentor *but* we might change it later or use 2 pairs (or aggregate the two control sites...)

note 2- same goes for cap_de_creus - no pair so i ommited but we should talk to the team and see what they think.

```{r}
spain<-dd_data %>% filter(country == "Spain")

spain<-gather(spain,"index","richness",22:24)

spain <- spain %>% mutate(pair = case_when(
  new_site_name == "formentera"~ "Formentera",
  new_site_name == "eivissa"~ "Formentera",
  new_site_name == "cabrera"~"Cabrera",
  new_site_name == "dragonera"~ "Cabrera",
  new_site_name == "cavalleria"~ "Cavalleria",
  new_site_name == "illa_de_l'aire"~ "Cavalleria",
  new_site_name == "medes"~ "Medes"
  
  ))

spain<-spain %>% drop_na(pair)


site_list[[4]]<-spain

spain$site<-spain$pair

spain <- spain %>% group_by(country,site,protection,enforcement,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

spain$pair <- spain$site

pair_list[[4]] <- spain

```

Italy

Need to decide how to treat those 3 sites...

  new_site_name == "carloforte"~"pair_2" , # control
  new_site_name == "asinara"~ "pair_2", # protected 3 & 2
  new_site_name == "tavolara"~ "pair_2", # protected 3
  
i did two different comparison with carloforta as the control for both MPAs  
  
  

```{r}
itl<-dd_data %>% filter(country == "Italy",data.origin != "Belmaker")


itl<-gather(itl,"index","richness",22:24)

itl <- itl %>% mutate(pair = case_when(
  new_site_name == "portofino"~ "Portofino",
  new_site_name == "genova"~ "Portofino",
  new_site_name == "torre_guaceto"~ "Torre guaceto",
  new_site_name == "otranto"~ "Torre guaceto",
  new_site_name == "asinara"~ "Asinara", 
  new_site_name == "carloforte"~"Asinara"
  ))

itl<-itl %>% drop_na(pair)

itl_2<-dd_data %>% filter(country == "Italy",new_site_name %in% c("tavolara","carloforte"))

itl_2<-gather(itl_2,"index","richness",22:24)

itl_2$pair = "Tavolara"

itl<-rbind(itl,itl_2)

site_list[[5]]<-itl
itl$site<-itl$pair

itl <- itl %>% group_by(country,site,protection,index) %>% summarise("mean_richness" = mean(richness),"se" = std.error(richness), "sd" = sd(richness))

itl$pair <- itl$site

pair_list[[5]] <- itl


```



```{r}
log_data_raw<-bind_rows(site_list)

rm(list=setdiff(ls(), "log_data_raw"))

plot_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\Pairwise"

labl<-c(dark = "Dark diversity",
        observed_richness = "Observed richness",
        total="Total species pool")

```

```{r}

log_data<-log_data_raw %>% select(pair,protection,index,richness) 

log_data$pair<-as.factor(log_data$pair)

log_data$pair<-str_to_title(log_data$pair)

# dark

d_ratio<-log_data %>% filter(index == "dark")

d_ratio = batch_calc_ES(dat = d_ratio, grouping =  pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase	 = "Not protected")

d_ratio<-d_ratio%>% mutate(pair = fct_reorder(pair,Est))

d_ratio$Index = "Dark diversity"

#observed

o_ratio<-log_data %>% filter(index == "observed_richness")

o_ratio = batch_calc_ES(dat = o_ratio, grouping = pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

o_ratio$Index<-"Observed richness"

# Total

t_ratio<-log_data %>% filter(index == "total")

t_ratio = batch_calc_ES(dat = t_ratio, grouping = pair, condition = protection, outcome = richness,scale = "count", ES = "LRRi",baseline_phase = "Not protected")

t_ratio$Index<-"Total richness"

#plot

ggplot(data = d_ratio,aes(x=Est,y=pair,color = Index))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 5)+
  geom_point(data = o_ratio,aes(x=Est,y= pair,color=Index),size = 5)+
  geom_point(data = t_ratio,aes(x=Est,y= pair,color=Index),size =5)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+
  geom_errorbar(data = d_ratio,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = o_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = t_ratio ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C","#ACD672"))+
  theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())

#ggsave("ratio_se_dark_obs_total.png",path = plot_dir,width = 6,height = 10)
#ggsave("ratio_se_dark_obs_total.pdf",path = plot_dir,width = 6,height = 7)

```

# Add metafor summary

```{r}
protection<-log_data_raw %>% select(pair,enforcement) %>% distinct(,.keep_all = T) %>% filter(as.numeric(as.character(enforcement))>2)


```


### Add meta-analyses summery from metafor


```{r}
metafor_data<-log_data_raw

metafor_data$protection<-ifelse(metafor_data$protection=="Protected","MPA","control")

```

## dark

### Set the data right

```{r}

metafor_dark<-metafor_data %>%
  filter(index =="dark") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 




```


### calculate the effect sizes

using the escalc function 

```{r}
effect_sizes_dark <- escalc(     # Function for calculating effect sizes
    "ROM",                       # Specify the effect size we want to calculate 
    m1i = mean_MPA,              # MPA mean richness  
    n1i = n_MPA,                 # MPA sample size
    sd1i = sd_MPA,               # MPA SD
    m2i = mean_control,          # control mean richness
    n2i = n_control,             # control sample size
    sd2i = sd_control,           # control site SD
    data = metafor_dark)         # data for  meta-analysis
  
# calculate the se

effect_sizes_dark$se = sqrt(effect_sizes_dark$vi)

# calculate the overall effect

dark_model <- rma(yi,                       # effect size from each row in database
                  vi,                       # measure of variance from each row in database
                  method = "REML",          # Specifies fixed effects model
                  slab = pair,              # prepares  labels for the forest plot 
                  data = effect_sizes_dark) # dataframe used for model 

dark_model

# calculate the overall effect only for enforcement = 3 

protected_dark <-effect_sizes_dark %>% filter(pair %in% protection$pair) 


pro_dark_model <- rma(yi,
                      vi, 
                      method = "REML", 
                      slab = pair, 
                      data = protected_dark)


pro_dark_model
```

### insert to data frame


```{r}

# summary of all MPAs 

d_summary<-d_ratio[1,c(1:4,7)]

d_summary$pair <- "Overall"

d_summary$SE<-dark_model$se

d_summary$Est<-dark_model$b[[1]]

d_summary$p_val<-dark_model[["pval"]]

# summary of highly enforced MPAs 

d_summary_pro<-d_ratio[1,c(1:4,7)]

d_summary_pro$pair <- "Protected"

d_summary_pro$SE<-pro_dark_model$se

d_summary_pro$Est<-pro_dark_model$b[[1]]

d_summary_pro$p_val<-pro_dark_model[["pval"]]

# join 

d_summary<-rbind(d_summary,d_summary_pro)
```

## Observed richness

```{r}
# set data

metafor_obs<-metafor_data %>%
  filter(index =="observed_richness") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 

# calculate effect size

effect_sizes_obs <- escalc( 
    "ROM",
    m1i = mean_MPA,                
    n1i = n_MPA,                 
    sd1i = sd_MPA,               
    m2i = mean_control,          
    n2i = n_control,           
    sd2i = sd_control,           
    data = metafor_obs)

#calculate overall effect

obs_model<- rma(yi, 
                vi, 
                method = "REML", 
                slab = pair,
                data = effect_sizes_obs) 

obs_model

# overall effect only for enforcment = 3

protected_obs <- effect_sizes_obs %>% filter(pair %in% protection$pair) 


pro_obs_model <- rma(yi, 
                     vi, 
                     method = "REML", 
                     slab = pair, 
                     data = protected_obs)

pro_obs_model

```

### inset to dataframe

```{r}

# all mpa 

o_summary<-o_ratio[1,c(1:4,7)]

o_summary$pair <- "Overall"

o_summary$SE<-obs_model$se

o_summary$Est<-obs_model$b[[1]]

o_summary$p_val<-obs_model[["pval"]]


# protected data

o_summary_pro<-o_ratio[1,c(1:4,7)]

o_summary_pro$pair <- "Protected"

o_summary_pro$SE<-pro_obs_model$se

o_summary_pro$Est<-pro_obs_model$b[[1]]

o_summary_pro$p_val<-pro_obs_model[["pval"]]

o_summary<-rbind(o_summary,o_summary_pro)
```

## Total overall

```{r}
# set the data

metafor_total<-metafor_data %>%
  filter(index =="total") %>%
  group_by(pair,protection) %>% 
  summarise("mean"= mean(richness),
            "sd"=sd(richness),
            "n" = n()) %>% 
  pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 

# calculate effect size

effect_sizes_total <- escalc( 
    "ROM",
    m1i = mean_MPA,               
    n1i = n_MPA,                 
    sd1i = sd_MPA,               
    m2i = mean_control,         
    n2i = n_control,             
    sd2i = sd_control,          
    data = metafor_total)



total_model<- rma(yi, 
                  vi, 
                  method = "REML",
                  slab = pair, 
                  data = effect_sizes_total) 

total_model

protected_total <-effect_sizes_total %>% filter(pair %in% protection$pair) 


pro_total_model <- rma(yi, 
                       vi, 
                       method = "REML", 
                       slab = pair, 
                       data = protected_total)

pro_total_model


```

### insrt to dataframe

```{r}

# all MPA

t_summary<-t_ratio[1,c(1:4,7)]

t_summary$pair <- "Overall"

t_summary$SE<-total_model$se

t_summary$Est<-total_model$b[[1]]

t_summary$p_val<-total_model[["pval"]]

# only highly protected MPAs


t_summary_pro<-t_ratio[1,c(1:4,7)]

t_summary_pro$pair <- "Protected"

t_summary_pro$SE<-pro_total_model$se

t_summary_pro$Est<-pro_total_model$b[[1]]

t_summary_pro$p_val<-pro_total_model[["pval"]]

t_summary<-rbind(t_summary,t_summary_pro)
```

## plot summraries


```{r}
ggplot(data = d_summary,aes(x=Est,y=fct_reorder(pair,Est),color = Index))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 6)+
  geom_point(data = o_summary,aes(x=Est,y= pair,color=Index),size = 6)+
  geom_point(data = t_summary,aes(x=Est,y= pair,color=Index),size = 6)+
  theme_classic()+
  ylab("MPA")+
  xlab("Log ratio")+xlim(-0.25,0.25)+
  geom_errorbar(data = d_summary,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = o_summary ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  geom_errorbar(data = t_summary ,aes(xmin= Est-SE, xmax=Est+SE), width=.2)+
  scale_color_manual(values=c('#7DCDE4',"#F2E96C","#ACD672"))+
  theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())
  

#ggsave("metafor_summary.png",path = plot_dir,width = 12,height = 8)
#ggsave("metafor_summary.pdf",path = plot_dir,width = 6,height = 3)





```


one table with all p-val

```{r}
all_model_summary<-rbind(d_summary,o_summary,t_summary)
```

# Completness

clean environment

```{r}
rm(list=setdiff(ls(), c("log_data","log_data_raw")))

completness_dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\plots\\completness"

```

Set the completeness data and calculate completeness per transect

```{r}
completeness<-log_data %>% group_by(pair,protection,index) %>% summarise("mean_richness"= mean(richness))


completeness<-spread(completeness,index,mean_richness)

completeness<-completeness %>% mutate("completness" = log(observed_richness/dark))


completeness_ratio<- completeness %>% select(pair,protection,completness) 

completeness_ratio<- spread(completeness_ratio,protection,completness)

 completeness_ratio<-completeness_ratio%>% 
  mutate("ratio"= `Protected`/`Not protected`) %>% 
  mutate("log_ratio" = log(ratio),"gap" = `Protected`-`Not protected`)

completeness_ratio<-completeness_ratio %>% mutate(pair = fct_reorder(pair,gap))

ggplot(completeness_ratio,aes(x=gap,y=fct_reorder(pair,gap)))+geom_point(size =5,color = "darkred")+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  ylab("MPA")+
  xlab("Completeness gap")+
  theme_classic()+theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())

#ggsave("completness gap.png",path = completness_dir,width = 12,height = 8)
#ggsave("completness gap.pdf",path = completness_dir,width = 12,height = 8)
```

```{r}

comp_transect <- log_data_raw %>% 
  select(pair, unique_trans_id, protection, index, richness) %>%
  filter(index != "total") %>%
  pivot_wider(names_from = index,
              values_from = richness) 

comp_transect$protection<-ifelse(comp_transect$protection=="Protected","MPA","control")

comp_transect$pair<-str_to_title(comp_transect$pair)

comp_transect<-comp_transect %>% mutate(completaness =log(observed_richness/dark))

comp_transect<-comp_transect %>% filter(completaness != Inf) %>% 
  group_by(pair,protection) %>%
  summarise(mean = mean(completaness),
            sd = sd(completaness),
            n = n())

comp_transect<-comp_transect%>% pivot_wider(names_from=protection, 
              values_from=c(mean, 
                            sd,
                            n)) 

comp_transect<-comp_transect%>%mutate(diffrance = mean_MPA - mean_control)

```

```{r}
effect_sizes_comp <- escalc( 
    "MD",
    m1i = mean_MPA,               
    n1i = n_MPA,                 
    sd1i = sd_MPA,               
    m2i = mean_control,         
    n2i = n_control,             
    sd2i = sd_control,
    var.names=c("yi","vi"),
    data = comp_transect)




forest(
  effect_sizes_comp$yi, # These are effect sizes from each row in database
  effect_sizes_comp$vi, # These are variances from each row in database
  annotate = T,            # Setting this to false prevents R from including CIs for each effect sizes 
  order="obs",             # set the order of the sites 
  slab = effect_sizes_comp$pair, # A individual effect size labeled 
  xlab = "Diferance", # Label for x-axis
  cex = .8,                    # Text side for study labels
  pch = 15,                    # shape of bars in forest plot
  cex.lab = 1                  # Size of x-axis label
)



a<-ggplot(effect_sizes_comp,aes(x=yi,y=fct_reorder(pair,yi)))+geom_point(size =5,color = "darkred")+geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  ylab("MPA")+
  xlab("Completeness difference")+ xlim(c(-0.6,0.6))+
  theme_classic()+theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())


comp_model<- rma(yi, 
                  vi, 
                  method = "REML",
                  slab = pair, 
                  data = effect_sizes_comp)


comp_model

c_summary<-data.frame(a=1)

c_summary$pair <- "Overall"

c_summary$ci_l<-comp_model$ci.lb
c_summary$ci_u<-comp_model$ci.ub
c_summary$Est<-comp_model$b[[1]]

c_summary$p_val<-comp_model[["pval"]]

c_summary$a<-NULL



b<-ggplot(data = c_summary,aes(x=Est,y=pair))+
  geom_vline(xintercept = 0,linetype = "dashed",size = 1)+
  geom_point(size = 6,color ="darkred")+
  theme_classic()+
  ylab("MPA")+
  xlab("Completeness difference")+xlim(-0.6,0.6)+
  geom_errorbar(data = c_summary,aes(xmin= ci_l, xmax=ci_u), width=.2,color ="darkred")+
  theme(axis.text.x = element_text(size=22),
        axis.text.y = element_text(size =22,angle = 0),
        axis.title.x = element_text(size=24),
        axis.title.y = element_text(size=24),
        legend.position = 'none',
        strip.text.x = element_text(size = 22),
        strip.background = element_blank())

```

```{r}
library(patchwork)

combined_plot <- a / b +
  plot_layout( heights  = c(15,3))

#ggsave("c_metafor.png",path = completness_dir,width = 10,height = 10)
#ggsave("c_metafor.pdf",path = completness_dir,width = 10,height = 10)


```

