---
title: "Clean 3 - protection and complexity models"
author: "TG"
date: "2024-01-11"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}
library(DarkDiv)
library(tidyverse)
library(vegan)
library(plotrix)
library(sjPlot)
library(sjmisc)
library(gridExtra)
library(lme4)
library(lmerTest)
library(MuMIn)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(patchwork)

select <- dplyr::select

dir<-"G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\final plots and models\\protection and complexity models"

```

# In this script:

1. Models of the realtionship between `Protection` and diversity componenets - `dark diversity`, `observed richness`, `species pool` and `completness`

2. Models of the realtionship between `Complexity` and diversity componenets - `dark diversity`, `observed richness`, `species pool` and `completness`


# Input files:

loaded from the `clean output data` folder

* `Clean_medata_dd_lis.Rdata` - dark diversity calculation for MeData, with Lisandro data 

* `Clean_complexity_dd.Rdata` - dark diversity calculation for the Israel complexity data 

# Outputs files:

**Models tables **

all saved in directory

*  `protection_model_results.xls` - tables for protection models

*  `curvature_model_results.xls` -  tables for Curvature complexity models

*  `vr_model_results.xls` -  tables for vertical relief complexity modeld

* `completenes_model_results` - All completeness models

**Plots** 

all saved in `protection and complexity models` folder

* `predictions_plot.pdf` - plot of protection + curvature models (Fig 3)

* `vr_predictions_plot.pdf`- plot of vertical relief models (Appendix 5)

* `completeness_prediction_plot`- plots of completeness (Appendix 2)


# Protection models

load `dd_data` data

```{r}
#load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\Clean_MeData_dd.Rdata")

load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\Clean_MeData_dd_lis.Rdata")

load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\pairs.Rdata")
```

Final data preparation:

New option:

1. keep in the models only for paired sites of no protection vs fully or partly protected

2. remove transect with less than 1 in the observed richness

3. treat enforcement as factor



```{r}
pairs$Cluster <- coalesce(pairs$Cluster, pairs$pair)

cluster<-pairs %>% select(-pair) %>% distinct() %>% rename(pair = Cluster)
```


```{r}
protction_model_data<-left_join(dd_data,cluster)

protction_model_data<-protction_model_data %>% drop_na(pair)

# which pairs i have in the data:

unique(protction_model_data$pair)

# how many pairs:

length(unique(protction_model_data$pair))

# which sites I have in the data

unique(protction_model_data$Site)

# How many sites:

length(unique(protction_model_data$Site))

#  richness <1

protction_model_data<-protction_model_data %>% filter(observed_richness >1)

# drop sites level

protction_model_data$Site<-droplevels(protction_model_data$Site)

# How many transects in each pair:  

summary.factor(protction_model_data$pair)

# convert to factor

protction_model_data$enforcement<-as.factor(protction_model_data$enforcement)

# how many from each enforcement level:

summary.factor(protction_model_data$enforcement)

```

## Dark diversity model

new option


```{r}
# run model

Dark <- glmmTMB(log(dark+1) ~ enforcement + (1|pair),
                  data = protction_model_data,
                  family = gaussian())

# model summary

summary(Dark) 

# residuals check

plot(simulateResiduals(Dark))

# prediction for the different enforcement levels

pred_dark = ggpredict(Dark,terms = 'enforcement')

# extract the p-value for the plot


p <-summary(Dark)

p0<-round(as.numeric(p[["coefficients"]][["cond"]][[10]]),3)
p2<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)
p3<-round(as.numeric(p[["coefficients"]][["cond"]][[12]]),3)

pval<-data.frame(enforcement = as.factor(c("0","2","3")),
                 p_val = c(p0,p2,p3))

# add significant signs

pval<- pval %>% mutate(sign = case_when(p_val > 0.1 ~ "",
                                        p_val > 0.05 & p_val <= 0.1 ~ "",
                                        p_val > 0.01 & p_val <= 0.05 ~ "*",
                                        p_val > 0.001 & p_val <= 0.01 ~ "**",
                                        p_val <= 0.001 ~ "***"))


plot_dark_pro <- ggplot(data = pred_dark, aes(x = x, y = predicted)) +
  geom_bar(stat = "identity", fill = "#7DCDE4", width = 0.7) +
  geom_errorbar(aes(ymin = predicted - std.error, ymax = predicted + std.error), 
                width = 0.2, color = "black", position = position_dodge(width = 0.7)) +
  xlab("Protection") +
  scale_x_discrete(labels = levels(protction_model_data$enforcement)) +
  geom_text(data = pval,
            aes(x = enforcement,y =  0.5, label = sign),
            size = 5)+
  ylab("Dark diversity") + theme_classic()

plot_dark_pro

rm(p,p0,p2,p3,pval,pred_dark)
```


old option

site is nested within Basin and modeled as random effect


```{r}
# # run model
# 
# Dark <- glmmTMB(log(dark+1) ~ enforcement + (1|Basin:Site),
#                   data = protction_model_data,
#                   family = gaussian())
# 
# # model summary
# 
# summary(Dark)
# 
# # residuals check
# 
# plot(simulateResiduals(Dark))
# 
# # prediction for the different enforcement levels
# 
# pred_dark = ggpredict(Dark,terms = c('enforcement[1:3 by=1]'))
# 
# # extract the p-value for the plot
# 
# p <-summary(Dark)
# 
# p<-round(as.numeric(p[["coefficients"]][["cond"]][[8]]),3)
# 
# subtitle<- paste0("p-value = ",p)
# 
# plot_dark_pro<-ggplot()+
#   geom_line(data = pred_dark,aes(x= x,y = predicted),size=1.5,color= "#7DCDE4")+
#   geom_ribbon(data = pred_dark,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#7DCDE4")+
#   xlab("Protection") +
#   scale_x_continuous(limits=c(1,3),breaks = c(1,2,3)) +
#   ggtitle(label ="Dark diversity",
#           subtitle = subtitle)+
#   ylab("Dark diversity")+theme_classic()
# 
# plot_dark_pro
# 
# rm(p,subtitle,pred_dark)

```


## Observed richness model

new version

```{r}

# run model

Observed <- glmmTMB(log(observed_richness) ~ enforcement + (1|pair),
                  data = protction_model_data,
                  family = gaussian())
# model summary

summary(Observed)

# plot Residuals

plot(simulateResiduals(Observed))

# predict the values for the differnt enforcment level

pred_observed <- ggpredict(Observed, terms = "enforcement")

# extract the p-value for the plot


p <-summary(Observed)

p0<-round(as.numeric(p[["coefficients"]][["cond"]][[10]]),3)
p2<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)
p3<-round(as.numeric(p[["coefficients"]][["cond"]][[12]]),3)

pval<-data.frame(enforcement = as.factor(c("0","2","3")),
                 p_val = c(p0,p2,p3))

# add significant signs

pval<- pval %>% mutate(sign = case_when(p_val > 0.1 ~ "",
                                        p_val > 0.05 & p_val <= 0.1 ~ "",
                                        p_val > 0.01 & p_val <= 0.05 ~ "*",
                                        p_val > 0.001 & p_val <= 0.01 ~ "**",
                                        p_val <= 0.001 ~ "***"))
# plot

plot_obsrved_pro <- ggplot(data = pred_observed, aes(x = x, y = predicted)) +
  geom_bar(stat = "identity", fill = "#F2E96C", width = 0.7) +
  geom_errorbar(aes(ymin = predicted - std.error, ymax = predicted + std.error), width = 0.2, color = "black", position = position_dodge(width = 0.7)) +
  xlab("Protection") +
  scale_x_discrete(labels = levels(protction_model_data$enforcement)) +
  geom_text(data = pval,
            aes(x = enforcement,y =  1, label = sign),
            size = 5)+
  ylab("Observerd diversity") + theme_classic()

plot_obsrved_pro

rm(p,p0,p2,p3,pval,pred_observed)
```



old version

```{r}
# 
# # run model
# 
# Observed <- glmmTMB(log(observed_richness) ~ enforcement + (1|Basin:Site),
#                   data = protction_model_data,
#                   family = gaussian())
# # model summary
# 
# summary(Observed)
# 
# # plot Residuals
# 
# plot(simulateResiduals(Observed))
# 
# # predict the values for the differnt enforcment level
# 
# pred_observed=ggpredict(Observed,terms = c('enforcement[1:3 by=1]'))
# 
# # get p-value
# 
# p <-summary(Observed)
# 
# p<-round(as.numeric(p[["coefficients"]][["cond"]][[8]]),3)
# 
# subtitle<- paste0("p-value = ",p)
# 
# # plot
# 
# plot_obsrved_pro<-ggplot()+
#   geom_line(data = pred_observed,aes(x= x,y = predicted),size=1.5,color= "#F2E96C")+
#   geom_ribbon(data = pred_observed,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#F2E96C")+
#   xlab("Protection") +
#   scale_x_continuous(limits=c(1,3),breaks = c(1,2,3)) +
#   ggtitle(label = "Observed richness",
#           subtitle = subtitle)+
#   ylab("Observed richness")+theme_classic()
# 
# plot_obsrved_pro
# 
# rm(p,subtitle,pred_observed)

```

## species pool model

new version

```{r}
# run model

Total <- glmmTMB(log(total) ~ enforcement + (1|pair),
                  data = protction_model_data,family = gaussian())
# summary

summary(Total)

# residuals

plot(simulateResiduals(Total))

# predicted values for the enforcment levels

pred_total=ggpredict(Total,terms = "enforcement")

# get p-values for the plot

p <-summary(Total)

p0 <- round(as.numeric(p[["coefficients"]][["cond"]][[10]]),3)
p2 <- round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)
p3 <- round(as.numeric(p[["coefficients"]][["cond"]][[12]]),3)

pval <-data.frame(enforcement = as.factor(c("0","2","3")),
                 p_val = c(p0,p2,p3))

# add significant signs

pval <- pval %>% mutate(sign = case_when(p_val > 0.1 ~ "",
                                        p_val > 0.05 & p_val <= 0.1 ~ "",
                                        p_val > 0.01 & p_val <= 0.05 ~ "*",
                                        p_val > 0.005 & p_val <= 0.01 ~ "**",
                                        p_val <= 0.005 ~ "***"))

# plot

plot_total_pro <- ggplot(data = pred_total, aes(x = x, y = predicted)) +
  geom_bar(stat = "identity", fill = "#ACD672", width = 0.7) +
  geom_errorbar(aes(ymin = predicted - std.error, ymax = predicted + std.error), width = 0.2, color = "black", position = position_dodge(width = 0.7)) +
  xlab("Protection") +
  scale_x_discrete(labels = levels(protction_model_data$enforcement)) +
  geom_text(data = pval,
            aes(x = enforcement,y =  1, label = sign),
            size = 5)+
  ylab("Species pool size") + theme_classic()

plot_total_pro

rm(p,p0,p2,p3,pval,pred_total)


```

old model

```{r}
# # run model
# 
# Total <- glmmTMB(log(total) ~ enforcement + (1|Basin:Site),
#                   data = protction_model_data,family = gaussian())
# # summary
# 
# summary(Total)
# 
# # residuals
# 
# plot(simulateResiduals(Total))
# 
# # predicted values for the enforcment levels
# 
# pred_total=ggpredict(Total,terms = c('enforcement[1:3 by=1]'))
# 
# # get p-values for the plot
# 
# p <-summary(Total)
# 
# p<-round(as.numeric(p[["coefficients"]][["cond"]][[8]]),3)
# 
# subtitle<- paste0("p-value = ",p)
# 
# # plot
# 
# plot_total_pro<-ggplot()+
#   geom_line(data = pred_total,aes(x= x,y = predicted),size=1.5,color= "#ACD672")+
#   geom_ribbon(data = pred_total,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#ACD672")+
#   xlab("Protection") +
#   scale_x_continuous(limits=c(1,3),breaks = c(1,2,3)) +
#   ggtitle(label = "Total species pool",
#           subtitle = subtitle)+
#   ylab("Species pool size")+theme_classic()
# 
# plot_total_pro
# 
# rm(p,subtitle,pred_total)

```

## save protection models tables


models summery tables:

**Note** - can be saved only to the directory, not to other folders

```{r}

#tab_model(Total, Observed, Dark, dv.labels = c("Species pool", "Observed richness", "Dark diversity"), file = "protection_model_results_new.xls")


```

## completeness - protection model

calculated as ln(observed richness/dark diversity)

Some transect have dark diversity = 0 and thus produce INF values, those transects was filterd out of this analyses.


```{r}

completeness<-protction_model_data %>%
  mutate(completeness = log(dark/observed_richness)) %>% filter(completeness != -Inf)


completeness_model <- glmmTMB(completeness ~ enforcement + (1|pair),
                  data = completeness,family = gaussian())

summary(completeness_model)

plot(simulateResiduals(completeness_model))

pred_completeness=ggpredict(completeness_model,terms = c('enforcement'))

# get p-values for the plot

p <-summary(completeness_model)

p0 <- round(as.numeric(p[["coefficients"]][["cond"]][[10]]),3)
p2 <- round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)
p3 <- round(as.numeric(p[["coefficients"]][["cond"]][[12]]),3)

pval <-data.frame(enforcement = as.factor(c("0","2","3")),
                 p_val = c(p0,p2,p3))

# add significant signs

pval <- pval %>% mutate(sign = case_when(p_val > 0.1 ~ "",
                                        p_val > 0.05 & p_val <= 0.1 ~ "",
                                        p_val > 0.01 & p_val <= 0.05 ~ "*",
                                        p_val > 0.001 & p_val <= 0.01 ~ "**",
                                        p_val <= 0.001 ~ "***"))

# plot

plot_completness_pro <- ggplot(data = pred_completeness, aes(x = x, y = predicted)) +
  geom_bar(stat = "identity", fill = "red", width = 0.7) +
  geom_errorbar(aes(ymin = predicted - std.error, ymax = predicted + std.error), width = 0.2, color = "black", position = position_dodge(width = 0.7)) +
  xlab("Protection") +
  scale_x_discrete(labels = levels(protction_model_data$enforcement)) +
  geom_text(data = pval,
            aes(x = enforcement,y =  -0.2, label = sign),
            size = 5)+
  ylab("Completeness") + theme_classic()+ geom_hline(yintercept=0)

plot_total_pro

rm(p,p0,p2,p3,pval,pred_completeness)

```


save completeness model table

```{r}
#tab_model(completeness_model,file = "Completeness_protection_model_new.xls")

```

# Complexity models

load complexity dark diversity data - `Clean_complexity_dd.Rdata` 

loaded as `comp_dd`

```{r}
load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\Clean_complexity_dd.Rdata")
```


### final data orgenaizing

Data preparation for model

```{r}
# remove transect with less than 2 species in the observed 

complexity_for_model<-comp_dd %>% filter(observed_richness >1)

# add column with completeness calculation 

complexity_for_model<-complexity_for_model %>% 
  mutate(completeness = log(observed_richness/dark))

# data for completeness models without INF values

complexity_comp <-complexity_for_model %>% filter(completeness != Inf)
```


first presented all the models for the curvature complexity index (Figure 3)

second presented all the models for the vertical relief complexity index (Appendix)

## Curvature index

#### Curvature dark diversity 

```{r}

# run model 

curv_dark<-glmmTMB(log(dark+1) ~  Curvature + protection + (1|site), data=complexity_for_model,family = gaussian())

# summary

summary(curv_dark)

# residuals plots

plot(simulateResiduals(curv_dark))

# predicted dark diversity for curvature values

dark_curv_predict =ggpredict(curv_dark,terms = c('Curvature[-2:4 by=0.5]'))

# add p-value for the plot

p <-summary(curv_dark)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot 

plot_dark_curv <-ggplot()+
  geom_line(data = dark_curv_predict,aes(x= x,y = predicted),size=1.5,color= "#7DCDE4")+
  geom_ribbon(data = dark_curv_predict,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#7DCDE4")+
  ggtitle(label = "Curvature dark diversity",
          subtitle = subtitle)+
  xlab("Curvature") +
  ylab("Dark diversity")+theme_classic()

plot_dark_curv

rm(p,subtitle,dark_curv_predict)

#ggsave("Curvature_dark.png",path = dir,width = plot_w,height = plot_h)
#ggsave("Curvature_dark.pdf",path = dir,width = plot_w,height = plot_h)

```

#### Curvature observed richness

```{r}

# run model

curv_observed<-glmmTMB(log(observed_richness) ~  Curvature + protection + (1|site), data=complexity_for_model,family = gaussian())

# summary

summary(curv_observed)

# residuals

plot(simulateResiduals(curv_observed))

# predicted richness

obs_curv_predict=ggpredict(curv_observed,terms = c('Curvature[-2:4 by=0.5]'))


# add p-value for the plot

p <-summary(curv_observed)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot

plot_observed_curv <- ggplot()+
  geom_line(data = obs_curv_predict,aes(x= x,y = predicted),size= 1.5,color= "#F2E96C")+
  geom_ribbon(data = obs_curv_predict,aes(x=x,ymin=conf.low,ymax=conf.high),alpha= 0.2,fill="#F2E96C")+
  ggtitle(label = "Curvature observed richness",
          subtitle = subtitle)+
  xlab("Curvature") + 
  ylab("Observed richness")+theme_classic()

plot_observed_curv

rm(p,subtitle,obs_curv_predict)

# ggsave("Curvature_obs.png",path = dir,width = plot_w,height = plot_h)
# ggsave("Curvature_obs.pdf",path = dir,width = plot_w,height = plot_h)
```

#### Curvature species pool


```{r}
# run model

curv_total <- glmmTMB(log(total) ~  Curvature + protection + (1|site), data=complexity_for_model,family = gaussian())

# summary

summary(curv_total)

# plot residuals

plot(simulateResiduals(curv_total))


# predicted species pool values

sppool_curv_pred =ggpredict(curv_total,terms = c('Curvature[-2:4 by=0.5]'))

# add p-value for the plot

p <-summary(curv_total)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

plot_total_curv<-ggplot()+
  geom_line(data = sppool_curv_pred,aes(x= x,y = predicted),size=1.5,color= "#ACD672")+
  geom_ribbon(data = sppool_curv_pred,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#ACD672")+  ggtitle(label = "Curvature species pool",
          subtitle = subtitle)+
  xlab("Curvature") +
  ylab("Species pool")+theme_classic()

plot_total_curv


rm(p,subtitle,sppool_curv_pred)


#ggsave("Curvature_total.png",path = plot_dir,width = 12,height = 7)
#ggsave("Curvature_total.pdf",path = plot_dir,width = 12,height = 7)
```

#### Curvature completness

```{r}
# run model

completness_curve<-glmmTMB(completeness ~  Curvature + protection+ (1|site), data = complexity_comp,family = gaussian())

# summary

summary(completness_curve)

# model residuals

plot(simulateResiduals(completness_curve))

# predict completeness

comp_curve_pred=ggpredict(completness_curve,terms = c('Curvature[-2:4 by=0.5]'))

# extract p-value for plot

p <-summary(completness_curve)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

#plot

plot_completeness_curv<-ggplot()+
  geom_line(data = comp_curve_pred,aes(x= x,y = predicted),size=1.5,color= "red")+
  geom_ribbon(data = comp_curve_pred,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="red")+
  xlab("Curvature") +
  ggtitle(label = "Curvature species pool",
          subtitle = subtitle)+
  ylab("Completeness")+theme_classic()


plot_completeness_curv

rm(p,subtitle,comp_curve_pred)

#ggsave("Curvature_comp.png",path = dir,width = plot_w,height = plot_h)
#ggsave("Curvature_comp.pdf",path = dir,width = plot_w,height = plot_h)
```

#### Save curvature models table

models summery tables:

**Note** - can be saved only to the directory, not to other folders

```{r}

# tab_model(curv_total, curv_observed, curv_dark, dv.labels = c("Species pool", "Observed richness", "Dark diversity"), file = "curvature_model_results.xls")

```

# Plot figure 3

combine all 6 plots of model predictions


```{r}

a<- plot_dark_curv
b<- plot_dark_pro
c<-plot_observed_curv
d<-plot_obsrved_pro
e<-plot_total_curv
f<-plot_total_pro

predictions_plot <- (a+b)/(c+d)/(e+f)+
  plot_layout(heights = rep(10), widths = rep(5))

predictions_plot
  
ggsave("predictions_plot_new.pdf",predictions_plot,path = dir, width = 10, height = 10)

rm(a,b,c,d,e,f)
```

## Vertical relief models

### vr dark diversity

```{r}

# model

vr_dark<-glmmTMB(log(dark+1) ~  VerRelief + protection+ (1|site), data=complexity_for_model,family = gaussian())

#summary

summary(vr_dark)

# plot mosel residuals

plot(simulateResiduals(vr_dark))

# predicted dark diversity values

predicted_dark_vr=ggpredict(vr_dark,terms = c('VerRelief[-2:4 by=0.5]'))

# extract p-value for plot

p <-summary(vr_dark)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot


plot_dark_vr <- ggplot()+
  theme_classic()+
  geom_line(data = predicted_dark_vr,aes(x= x,y = predicted),size=1.5,color= "#7DCDE4")+
  geom_ribbon(data = predicted_dark_vr,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#7DCDE4")+
  xlab("Vertical relief") +
  ylab("Dark diversity")+
  ggtitle(label = "VR dark diversity",
          subtitle = subtitle)

plot_dark_vr

rm(p,subtitle,predicted_dark_vr)

  
#ggsave("vr_dark.png",path = plot_dir,width = 12,height = 7)
#ggsave("vr_dark.pdf",path = plot_dir,width = 12,height = 7)
```

### vr observed richness


```{r}

# run model
vr_observed<-glmmTMB(log(observed_richness) ~  VerRelief + protection+ (1|site), data=complexity_for_model,family = gaussian())

# summary

summary(vr_observed)

# plot model residuals

plot(simulateResiduals(vr_observed))

# predict vr values

predict_obs_vr=ggpredict(vr_observed,terms = c('VerRelief[-2:4 by=0.5]'))

# extract p-value for plot

p <-summary(vr_observed)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot

plot_observed_vr<-ggplot()+
  theme_classic()+
  geom_line(data = predict_obs_vr,aes(x= x,y = predicted),size=1.5,color= "#F2E96C")+
  geom_ribbon(data = predict_obs_vr,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#F2E96C")+
  xlab("Vertical relief") +
  ylab("Observed richness")+
  ggtitle(label = "VR observed richness",
          subtitle = subtitle)

plot_observed_vr

rm(p,subtitle,predict_obs_vr)

#ggsave("vr_observed.png",path = plot_dir,width = 12,height = 7)
#ggsave("vr_observed.pdf",path = plot_dir,width = 12,height = 7)
```

### vr species pool


```{r}
# run model
vr_total<-glmmTMB(log(total) ~  VerRelief + protection + (1|site), data=complexity_for_model,family = gaussian())

# summary

summary(vr_total)

# plot model residuals

plot(simulateResiduals(vr_total))

# predicted species pool values

predicted_total_vr=ggpredict(vr_total,terms = c('VerRelief[-2:4 by=0.5]'))

# extract p-value for plot

p <-summary(vr_total)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot

plot_total_vr<-ggplot()+
  theme_classic()+
  geom_line(data = predicted_total_vr,aes(x= x,y = predicted),size=1.5,color= "#ACD672")+
  geom_ribbon(data = predicted_total_vr,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="#ACD672")+
  xlab("Vertical relief") +
  ylab("Total species pool")+
  ggtitle(label = "VR species pool",
          subtitle = subtitle)

plot_total_vr

rm(predicted_total_vr,p,subtitle)
#ggsave("vr_total.png",path = dir,width = 12,height = 7)
#ggsave("vr_total.pdf",path = dir,width = 12,height = 7)

```


### vr completness


```{r}
# model
completeness_vr<-glmmTMB(completeness ~  VerRelief + protection+ (1|site), data=complexity_comp,family = gaussian())

# summary

summary(completeness_vr)

# plot model residuals

plot(simulateResiduals(completeness_vr))

# predicted completeness values

predicted_comp_vr=ggpredict(completeness_vr,terms = c('VerRelief[-2:4 by=0.5]'))

# extract p-value for plot

p <-summary(completeness_vr)

p<-round(as.numeric(p[["coefficients"]][["cond"]][[11]]),3)

subtitle<- paste0("p-value = ",p)

# plot

plot_completeness_vr<-ggplot()+
  theme_classic()+
  geom_line(data = predicted_comp_vr,aes(x= x,y = predicted),size=1.5,color= "red")+
  geom_ribbon(data = predicted_comp_vr,aes(x=x,ymin=conf.low,ymax=conf.high),alpha=0.2,fill="red")+
  xlab("Vertical relief") +
  ylab("Completeness")+
  ggtitle(label = "VR completeness",
          subtitle = subtitle)

plot_completeness_vr

rm(predicted_comp_vr,p,subtitle)

#ggsave("vr_completness.png",path = completness_dir,width = 12,height = 7)
#ggsave("vr_completness.pdf",path = completness_dir,width = 12,height = 7)

```

### Save vr models table

models summery tables:

**Note** - can be saved only to the directory, not to other folders

```{r}

# tab_model(vr_total, vr_observed, vr_dark, dv.labels = c("Species pool", "Observed richness", "Dark diversity"), file = "vr_model_results.xls")

```


# Plot vertical relief appendix

```{r}
a<- plot_dark_vr
b<- plot_observed_vr
c<- plot_total_vr


vr_predictions_plot <- a/b/c +
  plot_layout(heights = rep(10), widths = rep(5))

vr_predictions_plot
  
ggsave("vr_predictions_plot.pdf",vr_predictions_plot,path = dir, width = 5, height = 10)

rm(a,b,c)

```

# Plot completness appendix


```{r}
a<- plot_completness_pro
b<- plot_completeness_curv
c<- plot_completeness_vr


completeness_prediction_plot <- a/b/c +
  plot_layout(heights = rep(10), widths = rep(5))

completeness_prediction_plot
  
ggsave("completeness_prediction_plot.pdf",completeness_prediction_plot,path = dir, width = 5, height = 10)

rm(a,b,c)

```

### Save completness models table

models summery tables:

**Note** - can be saved only to the directory, not to other folders

```{r}

# tab_model(completeness_model, completness_curve, completeness_vr, dv.labels = c("Protection", "Curvature", "Vertical relief"), file = "completenes_model_results_new.xls")

```


