---
title: "Clean 5 - Maps and summary tables"
author: "TG"
date: "2023-12-10"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r message=FALSE, warning=FALSE}
library(lubridate)
library(ggplot2)
library(stringr)
library(rfishbase)
library(Hmisc)
library(vegan)
library(mobr)
library(viridis)
library(plotrix)
library(sjmisc)
library(sf)
library(ggspatial)
library(dplyr)
library(tidyr)
library(leaflet)
library(rmapshaper)
library(patchwork)


dir <- "G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\final plots and models\\maps"

```

# In this script:

1. Mediterranean Sea map

2. Map for the 3 case sudies

3.


# Input files:

loaded from the `clean output data` folder

* `Clean_medata_dd.Rdata` - dark diversity calculation for MeData 

* `unique_coords` - dataframe with distinct survey locations


# Outputs files:

Maps are saved in the `maps` folder

* `combined_map` - plot of all maps

Tables are saved at directory folder

*`Transects table`- the number of transects conducted on each site  (Appendix 7)

*`Basin table` - the number of transects and sites in each Mediterranean Sea basin (Appendix 8)


```{r}
select <- dplyr::select

load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\unique_coords.Rdata")

load("G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\clean output data\\Clean_MeData_dd.Rdata")
```

# Part 1 - Maps

## load study site map

```{r}
med_seas <- st_read("Med_World_Seas_IHO_v3_MarineRegions/Medit_seas_IHO.shp")

```

### Banyuls

crop only the area of Banyuls MPA

```{r}

banyus_cropped <- st_crop(med_seas, xmin = 3.1, xmax = 3.18,
                                    ymin = 42.46, ymax = 42.53)

banyus<-unique_coords %>% filter(country=="France")

```

### Medes

crop only the area of Medes MPA


```{r}

Medes<-unique_coords %>% filter(new_site_name=="medes")

Medes_cropped <- st_crop(med_seas, xmin = 3.169, xmax = 3.26,
                                    ymin = 42.03, ymax = 42.11)
```

### Brijoni

crop only the area of Brijoni MPA

```{r}

brijuni<-unique_coords %>% filter(new_site_name=="brijuni")

brijuni_cropped <- st_crop(med_seas, xmin = 13.7, xmax = 13.85,
                                    ymin = 44.83, ymax = 44.95)

```

## Plot Mediterranean Sea map

```{r}
Med_map <- ggplot(unique_coords) +
  geom_sf(data = med_seas$geometry, mapping = aes(),fill= "azure",color = NA) + 
  geom_point(data =unique_coords,aes(x = lon, y = lat, color = enforcement,size = 1),position=position_jitter(h=0.06, w=0.06)) + 
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"))+
  theme(legend.position = "none",
        panel.grid.major = element_line(colour = "transparent"))+
        scale_color_manual("Protection level",values=c("#C42E1C","#D4B04A","#70A34E","#446441"))

Med_map

```

## Plots of case studies

find the center lat and lon for plotting porpuse

```{r}
b_bounds<-st_bbox(banyus_cropped$geometry)
c_bounds<-st_bbox(Medes_cropped$geometry)
d_bounds<-st_bbox(brijuni_cropped$geometry)

b_central_lat <- (b_bounds[[4]] + b_bounds[[2]]) / 2
b_central_lon <- (b_bounds[[3]] + b_bounds[[1]]) / 2

c_central_lat <- (c_bounds[[4]] + c_bounds[[2]]) / 2
c_central_lon <- (c_bounds[[3]] + c_bounds[[1]]) / 2

d_central_lat <- (d_bounds[[4]] + d_bounds[[2]]) / 2
d_central_lon <- (d_bounds[[3]] + d_bounds[[1]]) / 2

```

plot the case-sutdy sites

```{r}

banyus<-ggplot() +
  geom_sf(data = banyus_cropped,fill= "azure",color = NA)+ 
  geom_point(data =banyus,aes(x = lon, y = lat, color = enforcement),size = 3)+
  theme_bw()+
  theme(panel.background = element_rect(fill = "grey90",color = "black"),
       legend.position = "none",axis.title =  element_blank(),
        panel.grid.major = element_line(colour = "transparent"))+
        scale_color_manual("Protection level",values=c("#C42E1C","#446441"))+
  annotation_scale(location = "bl", width_hint = 0.4)+
    scale_x_continuous(expand = c(0, 0), breaks = round(b_central_lon,2)) +
  scale_y_continuous(expand = c(0, 0), breaks = round(b_central_lat,2))
    
banyus


Medes <-ggplot() +
  geom_sf(data = Medes_cropped,fill= "azure",color = NA)+ 
  geom_point(data =Medes,aes(x = lon, y = lat, color = enforcement),size = 3)+
  theme_bw()+
  theme(panel.background = element_rect(fill = "grey90",color = "black"),
        legend.position = "none",axis.title =  element_blank(),
        panel.grid.major = element_line(colour = "transparent"))+
        scale_color_manual("Protection level",values=c("#C42E1C","#446441"))+
  annotation_scale(location = "bl", width_hint = 0.4)+
      scale_x_continuous(expand = c(0, 0), breaks = round(c_central_lon,2)) +
  scale_y_continuous(expand = c(0, 0), breaks = round(c_central_lat,2))

Medes


brijuni<-ggplot() +
  geom_sf(data = brijuni_cropped,fill= "azure",color = NA)+ 
  geom_point(data =brijuni,aes(x = lon, y = lat, color = enforcement),size =3)+
  theme_bw()+
  theme(panel.background = element_rect(fill = "grey90",color = "black"),
       legend.position = "none",
       axis.title =  element_blank(),
        panel.grid.major = element_line(colour = "transparent"))+
        scale_color_manual("Protection level",values=c("#C42E1C","#446441"))+
  annotation_scale(location = "bl", width_hint = 0.4)+
      scale_x_continuous(expand = c(0, 0), breaks = round(d_central_lon,2)) +
  scale_y_continuous(expand = c(0, 0), breaks = round(d_central_lat,2))

brijuni

```

## Plot all Maps and save

```{r}
combined_map <- Med_map / (banyus + Medes + brijuni) +
  plot_layout(ncol = 1,nrow = 3,widths = c(14,7,7,7))
   
# Print the combined plot

combined_map

#ggsave(combined_map,filename = "combined_map.pdf",path = dir , width=10, height=10)
```

# Part 2 - tables and summary statistics


```{r}
rm(list=setdiff(ls(), "dd_data"))

dir<- "G:\\My Drive\\PHD\\chapter 3 MPA dd\\Meddata-DD\\final plots and models\\transect tables"

```


# Transect table


```{r}
dd_data$Site <- sub("_", " ", dd_data$Site)
dd_data$Site<-str_to_title(dd_data$Site)
```

```{r}
Transects_table <- dd_data %>%
  group_by(country,Site, protection) %>%
  summarise("Transect number" = n(),
            "Mean observed richness" = round(mean(observed_richness),1),
            "Mean dark diversity" = round(mean(dark),1)) 

Transects_table$country<-as.character(Transects_table$country)  

Transects_table<-Transects_table %>% arrange(country,Site,protection)

# write.csv(Transects_table,paste0(dir,"\\Transects_table.csv"))
```



## Basin table


```{r}
Basin_table <- dd_data %>% 
  group_by(Basin) %>%
  summarise("Number of sites"= n_distinct(Site,protection),
            "Number of transects" = n_distinct(unique_trans_id)) %>%
  arrange(Basin)


#write.csv(Basin_table,paste0(dir,"\\Basin_table.csv"))

```


## Descriptive stats used in the manuscript

```{r echo=FALSE, results='asis'}

cat("* Overall transects conducted - ", nrow(dd_data), "\n")

cat("* Overall transects in fully protected area - ", nrow(filter(dd_data,enforcement == 3)), "\n")

cat("* Overall transects in partly protected area - ", nrow(filter(dd_data,enforcement == 2)), "\n")

cat("* Overall transects in minimally protected area - ", nrow(filter(dd_data,enforcement == 1)), "\n")

cat("* Overall transects in not protected area - ", nrow(filter(dd_data,enforcement == 0)), "\n")

cat("* max sampling effort per site - ", max(Transects_table$`Transect number`), "\n")

cat("* min sampling effort per site - ", min(Transects_table$`Transect number`), "\n")

```
  