---
title: "Data acquaintance for Mediterranean dataset (MEData)"
author: "Shira Salingré"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
library(sf)
library(plotly)
library(tidyverse)
library(ggspatial)
```

# Introduction
Medata is a dataset of underwater fish census surveys conducted along the Mediterranean sea. The dataset includes environmental variables, as well as conservation information regarding MPAs.

This is a collaboration of [Belmaker Lab](https://belmaker.weebly.com/) (Tel Aviv University, Israel), [Ernesto Azzuro](https://www.researchgate.net/profile/Ernesto_Azzurro) (ISPRA, Italy), [Joachim Claudet](http://www.joachimclaudet.com/) (CNRS, France) and [Enric Sala](https://www.researchgate.net/profile/Enric_Sala).


```{r setup data}
med_raw <- read_csv("data/uvc_data_19022020.csv", col_types = cols(depth = "d"))

# locations will be a subset of unique sites: sites that have unique longitude and latitude
locations <- med_raw %>% 
  dplyr::select(lon, lat, country, site) %>% 
  dplyr::distinct(lon, lat, .keep_all = TRUE)
count(locations)
```

# Variables

| variable        | type  | comments                             |
| :-------------- | :---: | :------------------------------------|
|  `data.origin`  | *fct* | azz_asi = Azzuro; azz_malta = Azzuro; Belmaker = Belmaker; Claudet = Claudet; Sala - PEW = Sala |
| `country`       | *fct* | Country where `site` is located |
| `sites`         | *fct* | site names   |
| `lon`           | *num* | longitude (degrees); approximate for Linosa, Kornati|
| `lat`           | *num* | latitude (degrees); approximate for Linosa, Kornati |
| `trans`         | *int* | transect number |
| `species`       | *fct* | species scientific name (*Genus.species*); some species are recognised to Genus (spp)  or Family (-dae suffix) level |
| `sp.length`     | *num* | length of fish in cm |
| `sp.n`          | *int* | fish count - how many individuals of this species were observed? |
| `season`        | *fct* | levels: `Autumn`, `Spring`, `Summer` - The season in which the survey was conducted |
| `protection`    | *fct* | inside MPA = `YES`; outside MPA = `NO` |
| `enforcement`   | *int* | type of enforcement. 3 levels: `1` = Minimal protection; `2` = Medium protection; `3` = Fully protected. Complements `protection` |
| `total.mpa.ha`  | *num* | total MPA area in hactare; [Source for Ustica](http://www.parks.it/indice/RM/Eindex.php); [Source for Marettimo](http://www.ampisoleegadi.it/?idx=1620) |
| `size.notake`   | *num* | area size of no-take zone (in ha); [Source for Ustica](http://www.mmmpa.eu/UST-MPA.asp) |
| `yr.creation`   | *int* | MPA establishment year |
| `age.reserve.yr`| *int* | age of the MPA in years (corresponds to `yr.creation`) |
| `depth`         | *num* | depth of survey in metres (where applicable) |
| `a`             | *num* | species length-weight relationship constant |
| `b`             | *num* | species length-weight relationship constant |
| `tmean`         | *num* | annual mean temperature (Source: [Bio-ORACLE](http://www.bio-oracle.org/index.php)) |
| `trange`        | *num* | annual temperature range (Source: [Bio-ORACLE](http://www.bio-oracle.org/index.php)) |
| `sal_mean`      | *num* | annual mean salinity (Source: [Bio-ORACLE](http://www.bio-oracle.org/index.php)) |
| `pp_mean`       | *num* | annual mean primary productivity (Source: [Bio-ORACLE](http://www.bio-oracle.org/index.php)) |
| `pp_range`      | *num* | annual primary production range (Source: [Bio-ORACLE](http://www.bio-oracle.org/index.php)) |

# Overview

```{r}
summary(med_raw) # examine raw data and check for errors.

```

## Countries

Data was collected from 8 countries:

```{r review the dataset}
med_raw %>% distinct(country) %>% arrange(.$country) %>% print(n = Inf) # full country list

```

## Sites

Data was collected from 224 sites
```{r include=FALSE}
sites <- med_raw %>% distinct(site) %>% arrange(.$site)
count(sites)

```

## Species

Dataset withholds 118 species
```{r}
spp <- med_raw %>% distinct(species) %>% arrange(.$species)
count(spp)
spp %>% filter(!c(grepl("dae", species) | grepl(".spp", species))) %>% print(n = Inf)

```
**NOTE**: There are 123 taxa in the dataset but some are not species but genus (Atherina.spp, Symphodus.spp) or family (Labridae, Clupeidae, Belonidae).

# Methods

## Data collection
Visual fish census surveys took place in multiple locations along the Mediterranean Sea (figure 1) by teams of skilled SCUBA divers. Locations were comprised of sites within MPAs with varying size, age and enforcement level, and unprotected sites which are adjacent to these MPAs. To date, the database consists of approximately 42,0000 observations of 118 species of fish and includes mostly abundance data (except *Asinara_add* site which contains presence-absence data).

```{r locations map, echo=FALSE}
med_map <- sf::st_read("Med_World_Seas_IHO_v3_MarineRegions/Medit_seas_IHO.shp", quiet = TRUE)

sites_plot <- ggplot(med_raw) +
  geom_sf(data = med_map$geometry, mapping = aes()) +
  geom_point(aes(x = lon, y = lat, fill = tmean), colour = "black", pch = 21,
             size = 5, alpha = 0.7) +
  scale_fill_distiller(palette = "RdBu") +
  ggtitle(label = "Mediterranean fish surveys", subtitle = "Locations and temperatures") +
  xlab("Longitude") + 
  ylab("Latitude") +
  labs(fill = "Mean Annual Temp")

sites_plot

```
Figure 1. Survey sites and temperatures (in degrees celsius).

## Data acquired from other resources

**Environmental data** (temperature, salinity and primary production) were acquired from [Bio-ORACLE](http://http://www.bio-oracle.org) using `sdmpredictors` package.

__Layers__:

- tmean = "BO_sstmean"
- trange = "BO_sstrange"
- sal_mean = "BO2_salinitymean_ss"
- pp_mean = "BO2_ppmean_ss"
- pp_range = "BO2_pprange_ss"

(See "~/R/bio-oracle extraction code.R" for full extraction code).


## Data wrangling

* Linosa and kornati did not have specific coordinates, therefore, an approximate location (lat-lon) was attached to it. If your analysis requires fine-detail for the location, you might want to omit these locations.

* Data from source `ASINARA_add` site are presence-absence only.

### Transformations
Many analyses require species matrix where each row is a site and each column is a species.

Here's the code to create such matrix with this dataset:

```{r create species matrix}
# Create a tibble of metadata
med_meta <- med_raw %>% 
  distinct(country, site, lon, lat, season, protection, enforcement, total.mpa.ha, size.notake, 
           yr.creation, age.reserve.yr, tmean, trange, sal_mean, pp_mean, pp_range)

# Create a species matrix with summation of each species in each site (abundance)
med_mat <- med_raw %>% 
  group_by(site, lon, lat, species) %>% # Sites and species (with coordinate-locations) only
  summarise(n = sum(sp.n)) %>% 
  spread(species, n, fill = 0)

# Put the 2 data sets together (species and metadata)
full_med_mat <- left_join(med_meta, med_mat, by = c("site", "lon", "lat"))

# Convert the abundance data to pseudo presence-absence
pres_abs_mat <- full_med_mat
pres_abs_mat[17:ncol(pres_abs_mat)] <- ifelse(pres_abs_mat[17:ncol(pres_abs_mat)] > 0, 1, 0) # col 17 is the 1st species


```


